### 2017-03-06 workbooks/sparse_hawkes.Rmd
workbook import from private research repo

---
title: "Sparse hawkes"
author: "Dan MacKinlay"
date: "6 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-03-06 workbooks/sparse_hawkes.R

@@ -1,0 +1,4 @@
+require("glmnet")
+self_basis_expand = function(x, filters) {
+  
+}

### 2017-03-06 workbooks/tsglmcrosscheck.R

@@ -1,0 +1,6 @@
+library("tscount")
+library("feather")
+
+
+
+arma3=tsglm.sim(1000, param=list(pas))

### 2017-03-06 setup.py
hawkes_process->branching_process

@@ -4,7 +4,7 @@ except ImportError:
     from distutils.core import setup
 
 long_description = """
-hawkes_process
+branching_process
 """
 
 config = dict(
@@ -21,12 +21,12 @@ config = dict(
         'seaborn',
         'pandas',
     ],
-    packages=['hawkes_process'],
+    packages=['branching_process'],
     scripts=[],
-    name='hawkes_process',
+    name='branching_process',
     # # see https://python-packaging.readthedocs.io/en/latest/non-code-files.html
     # package_data=dict(
-    #     hawkes_process= ['datasets'],
+    #     branching_process= ['datasets'],
     # ),
     # include_package_data=True
 )

### 2017-03-06 branching_process/cts/__init__.py
integrate influence kernels

@@ -1,3 +1,3 @@
 from __future__ import absolute_import
 
-from .fit import Continuous
+from .fit import ContinuousExact

### 2017-03-06 branching_process/cts/influence.py

@@ -1,12 +1,13 @@
+have_autograd = False
+
 try:
     import autograd
-    import autograd.numpy as np
-    from autograd.numpy import sqrt, pi
     have_autograd = True
+    import autograd.numpy as np
+    import autograd.scipy as sp
 except ImportError as e:
     import numpy as np
-    have_autograd = False
-    from np import sqrt, pi
+    import scipy as sp
 
 
 class InfluenceKernel(object):
@@ -28,7 +29,7 @@ class UniModalInfluenceKernel(InfluenceKernel):
 
     def majorant(self, t, tau, *args, **kwargs):
         mode = self.mode(tau)
-        peak = self(mode)
+        peak = self(mode, tau)
         return np.choose(
             t > mode,
             [
@@ -43,19 +44,32 @@ class ExpKernel(InfluenceKernel):
         theta = 1.0 / tau
         return theta * np.exp(-t * theta) * (t >= 0)
 
+    def integrate(self, t, tau, *args, **kwargs):
+        theta = 1.0 / tau
+        return 1-np.exp(-t * theta) * (t >= 0)
+
 
 class MaxwellKernel(UniModalInfluenceKernel):
     """
+    http://mathworld.wolfram.com/MaxwellDistribution.html
     I think I could just use ``scipy.stats.maxwell``?
+    That seems not to be autograd differentiable.
     """
     def __call__(self, t, tau, *args, **kwargs):
         t2 = np.square(t)
-        return sqrt(2.0/pi) * t2 * np.exp(
+        return np.sqrt(2.0/np.pi) * t2 * np.exp(
             -t2 / (2 * tau**2)
         )/(tau**3)
 
     def mode(self, tau, *args, **kwargs):
-        return sqrt(2) * tau
+        return np.sqrt(2) * tau
+
+    def integrate(self, t, tau, *args, **kwargs):
+        return sp.special.erf(
+            t / (np.sqrt(2)*tau)
+        ) - t * np.sqrt(2.0/np.pi) / tau * np.exp(
+            -np.square(t)/(2 * np.square(tau))
+        )
 
 
 class MultiKernel(object):
@@ -64,7 +78,7 @@ class MultiKernel(object):
             n_kernels=1,
             kernel=MaxwellKernel(),
             *args, **kwargs):
-        self.kernel=kernel
+        self.kernel = kernel
 
     def majorant(self, t, tau=None, kappa=None, *args, **kwargs):
         return self(t, tau=tau, kappa=kappa, *args, **kwargs)
[…]

### 2017-03-06 branching_process/cts/influence.py
MultiKernel refix

@@ -80,28 +80,68 @@ class MultiKernel(object):
             *args, **kwargs):
         self.kernel = kernel
 
-    def majorant(self, t, tau=None, kappa=None, *args, **kwargs):
-        return self(t, tau=tau, kappa=kappa, *args, **kwargs)
+    def majorant(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
+        return self.majorant(t, tau=tau, kappa=kappa, *args, **kwargs)
 
-    def __call__(self, t, tau=None, kappa=None, *args, **kwargs):
+    def __call__(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
         pass
 
-    def integrate(self, t, tau=None, kappa=None, *args, **kwargs):
+    def integrate(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
         pass
 
-    def majorant_all(self, t, tau=None, kappa=None, *args, **kwargs):
+    def majorant_each(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
         out = np.zeros(self.n_kernels)
         if np.isscalar(kappa):
             kappa = np.ones(self.n_kernels) * kappa
         for i in range(self.n_kernels):
-            out[i] = self.kernel(t, tau=tau[i], kappa=kappa)
+            out[i] = self.kernel.majorant(t, tau=tau[i], kappa=kappa[i])
         return out
 
-    def call_all(self, t, tau=None, kappa=None, *args, **kwargs):
-        pass
+    def call_each(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
+        out = np.zeros(self.n_kernels)
+        if np.isscalar(kappa):
+            kappa = np.ones(self.n_kernels) * kappa
+        for i in range(self.n_kernels):
+            out[i] = self.kernel(t, tau=tau[i], kappa=kappa[i])
+        return out
 
-    def integrate_all(self, t, tau=None, kappa=None, *args, **kwargs):
-        pass
+    def integrate_each(
+            self,
+            t,
+            tau=None,
+            kappa=None,
+            *args, **kwargs):
+        out = np.zeros(self.n_kernels)
+        if np.isscalar(kappa):
+            kappa = np.ones(self.n_kernels) * kappa
+        for i in range(self.n_kernels):
[…]

### 2017-03-06 branching_process/cts/influence.py
simpler fixed kernel

@@ -144,29 +144,30 @@ class MultiKernel(object):
         return out
 
 
-# replace with a wrapper function? yes.
 class FixedKernel(InfluenceKernel):
+    """
+    if you want some arguments to be fixed, this wrapper will do it.
+    """
     def __init__(
             self,
             kernel,
-            tau=1.0,
-            kappa=1.0,
+            *args,
+            **fixed_args
             ):
-        self.tau = tau
-        self.kappa = kappa
         self.kernel = kernel
+        self.fixed_args = fixed_args
 
     def majorant(self, t, *args, **kwargs):
         return self.kernel.majorant(
-            t, tau=self.tau, kappa=self.kappa,
-            *args, **kwargs)
+            t,
+            *args, **kwargs, **self.fixed_args)
 
     def __call__(self, t, *args, **kwargs):
         return self.kernel(
-            t, tau=self.tau, kappa=self.kappa,
-            *args, **kwargs)
+            t,
+            *args, **kwargs, **self.fixed_args)
 
     def integrate(self, t, *args, **kwargs):
         return self.kernel.integrate(
-            t, tau=self.tau, kappa=self.kappa,
-            *args, **kwargs)
+            t,
+            *args, **kwargs, **self.fixed_args)

### 2017-03-06 branching_process/cts/influence.py
tidy up kernel implementation

@@ -19,7 +19,7 @@ class InfluenceKernel(object):
         raise NotImplementedError()
 
     def integrate(self, t, tau, *args, **kwargs):
-        # some kind of quadrature?
+        # some kind of numerical quadrature fallback?
         raise NotImplementedError()
 
 
@@ -72,7 +72,7 @@ class MaxwellKernel(UniModalInfluenceKernel):
         )
 
 
-class MultiKernel(object):
+class MultiKernel(InfluenceKernel):
     def __init__(
             self,
             n_kernels=1,
@@ -160,14 +160,35 @@ class FixedKernel(InfluenceKernel):
     def majorant(self, t, *args, **kwargs):
         return self.kernel.majorant(
             t,
-            *args, **kwargs, **self.fixed_args)
+            *args, **kwargs,
+            **self.fixed_args)
 
     def __call__(self, t, *args, **kwargs):
         return self.kernel(
             t,
-            *args, **kwargs, **self.fixed_args)
+            *args, **kwargs,
+            **self.fixed_args)
 
     def integrate(self, t, *args, **kwargs):
         return self.kernel.integrate(
             t,
-            *args, **kwargs, **self.fixed_args)
+            *args, **kwargs,
+            **self.fixed_args)
+
+    def majorant_each(self, t, *args, **kwargs):
+        return self.kernel.majorant_each(
+            t,
+            *args, **kwargs,
+            **self.fixed_args)
+
+    def call_each(self, t, *args, **kwargs):
+        return self.kernel.call_each(
+            t,
+            *args, **kwargs,
+            **self.fixed_args)
+
+    def integrate_each(self, t, *args, **kwargs):
+        return self.kernel.integrate_each(
+            t,
+            *args, **kwargs,
+            **self.fixed_args)

### 2017-03-06 branching_process/cts/influence.py
generic kernel

@@ -144,6 +144,42 @@ class MultiKernel(InfluenceKernel):
         return out
 
 
+class GenericKernel(InfluenceKernel):
+    """
+    Construct a kernel from some functions
+    """
+    def __init__(
+            self,
+            kernel,
+            majorant=None,
+            integral=None,
+            *args,
+            **fixed_args
+            ):
+        self._kernel = kernel
+        self._majorant = majorant if majorant is not None else kernel
+        self._integral = integral
+        self._fixed_args = fixed_args
+
+    def majorant(self, t, *args, **kwargs):
+        return self.majorant(
+            t,
+            *args, **kwargs,
+            **self._fixed_args)
+
+    def __call__(self, t, *args, **kwargs):
+        return self._kernel(
+            t,
+            *args, **kwargs,
+            **self._fixed_args)
+
+    def integrate(self, t, *args, **kwargs):
+        return self._integrate(
+            t,
+            *args, **kwargs,
+            **self._fixed_args)
+
+
 class FixedKernel(InfluenceKernel):
     """
     if you want some arguments to be fixed, this wrapper will do it.

### 2017-03-06 branching_process/cts/influence.py
inheritance design hell

@@ -12,41 +12,48 @@ except ImportError as e:
 
 class InfluenceKernel(object):
 
-    def majorant(self, t, tau, *args, **kwargs):
-        return self(t, tau=tau, *args, **kwargs)
+    def __init__(self, *args, **fixed_args):
+        self._fixed_args = fixed_args
+        super(InfluenceKernel, self).__init__(*args)
 
-    def __call__(self, t, tau, *args, **kwargs):
-        raise NotImplementedError()
+    def majorant(self, t, *args, **kwargs):
+        return self._majorant(t, *args, **kwargs, **self._fixed_args)
 
-    def integrate(self, t, tau, *args, **kwargs):
+    def __call__(self, t, *args, **kwargs):
+        return self._call(t, *args, **kwargs, **self._fixed_args)
+
+    def integrate(self, t, *args, **kwargs):
         # some kind of numerical quadrature fallback?
-        raise NotImplementedError()
+        return self._integrate(t, *args, **kwargs, **self._fixed_args)
 
 
 class UniModalInfluenceKernel(InfluenceKernel):
     def mode(self, tau, *args, **kwargs):
         return 0
 
-    def majorant(self, t, tau, *args, **kwargs):
-        mode = self.mode(tau)
-        peak = self(mode, tau)
+    def mode(self, tau, *args, **kwargs):
+        return 0
+
+    def _majorant(self, t, *args, **kwargs):
+        mode = self.mode(*args, **kwargs, **self._fixed_args)
+        peak = self(*args, **kwargs, **self._fixed_args)
         return np.choose(
             t > mode,
             [
                 peak,
-                self(t, tau, *args, **kwargs)
+                self(t, *args, **kwargs, **self._fixed_args)
             ]
         )
 
 
 class ExpKernel(InfluenceKernel):
-    def __call__(self, t, tau, *args, **kwargs):
+    def _call(self, t, tau, *args, **kwargs):
         theta = 1.0 / tau
         return theta * np.exp(-t * theta) * (t >= 0)
 
-    def integrate(self, t, tau, *args, **kwargs):
+    def _integrate(self, t, tau, *args, **kwargs):
         theta = 1.0 / tau
-        return 1-np.exp(-t * theta) * (t >= 0)
+        return 1 - np.exp(-t * theta) * (t >= 0)
 
 
 class MaxwellKernel(UniModalInfluenceKernel):
[…]

### 2017-03-06 branching_process/cts/influence.py
fixed params available verywher

@@ -17,7 +17,9 @@ class InfluenceKernel(object):
         super(InfluenceKernel, self).__init__(*args)
 
     def majorant(self, t, *args, **kwargs):
-        return self._majorant(t, *args, **kwargs, **self._fixed_args)
+        return getattr(
+            self, '_majorant', self._call
+        )(t, *args, **kwargs, **self._fixed_args)
 
     def __call__(self, t, *args, **kwargs):
         return self._call(t, *args, **kwargs, **self._fixed_args)
@@ -28,20 +30,21 @@ class InfluenceKernel(object):
 
 
 class UniModalInfluenceKernel(InfluenceKernel):
-    def mode(self, tau, *args, **kwargs):
-        return 0
+    def mode(self, *args, **kwargs):
+        return self._mode(**kwargs, **self._fixed_args)
 
-    def mode(self, tau, *args, **kwargs):
+    def _mode(self, *args, **kwargs):
         return 0
 
     def _majorant(self, t, *args, **kwargs):
-        mode = self.mode(*args, **kwargs, **self._fixed_args)
-        peak = self(*args, **kwargs, **self._fixed_args)
+        mode = self.mode(*args, **kwargs)
+        peak = self(mode, *args, **kwargs)
+        print('umk', mode, peak)
         return np.choose(
             t > mode,
             [
                 peak,
-                self(t, *args, **kwargs, **self._fixed_args)
+                self(t, *args, **kwargs)
             ]
         )
 
@@ -68,7 +71,7 @@ class MaxwellKernel(UniModalInfluenceKernel):
             -t2 / (2 * tau**2)
         )/(tau**3)
 
-    def mode(self, tau, *args, **kwargs):
+    def _mode(self, tau, *args, **kwargs):
         return np.sqrt(2) * tau
 
     def _integrate(self, t, tau, *args, **kwargs):
@@ -88,13 +91,24 @@ class MultiKernel(InfluenceKernel):
         self.kernel = kernel
         super(MultiKernel, self).__init__(*args, **fixed_args)
 
+    def _fixed_args_each(self):
+        for i in range(self.n_kernels):
+            yield dict([
+                (key, val[i])
+                for key, val
+                in self._fixed_args.iteritems()
+            ])
+
     def majorant(
             self,
             t,
             tau=None,
             kappa=None,
             *args, **kwargs):
-        return self.majorant(t, tau=tau, kappa=kappa, *args, **kwargs)
+
+        return np.sum(
[…]

### 2017-03-06 branching_process/cts/influence.py
alright this is a half-arsed plan but it kinda works

@@ -17,16 +17,22 @@ class InfluenceKernel(object):
         super(InfluenceKernel, self).__init__(*args)
 
     def majorant(self, t, *args, **kwargs):
-        return getattr(
-            self, '_majorant', self._call
-        )(t, *args, **kwargs, **self._fixed_args)
+        return np.squeeze(
+            getattr(
+                self, '_majorant', self._call
+            )(t, *args, **kwargs, **self._fixed_args)
+        )
 
     def __call__(self, t, *args, **kwargs):
-        return self._call(t, *args, **kwargs, **self._fixed_args)
+        return np.squeeze(
+            self._call(t, *args, **kwargs, **self._fixed_args)
+        )
 
     def integrate(self, t, *args, **kwargs):
         # some kind of numerical quadrature fallback?
-        return self._integrate(t, *args, **kwargs, **self._fixed_args)
+        return np.squeeze(
+            self._integrate(t, *args, **kwargs, **self._fixed_args)
+        )
 
 
 class UniModalInfluenceKernel(InfluenceKernel):
@@ -51,10 +57,14 @@ class UniModalInfluenceKernel(InfluenceKernel):
 
 class ExpKernel(InfluenceKernel):
     def _call(self, t, tau, *args, **kwargs):
+        t = np.asarray(t).reshape(1, -1)
+        tau = np.asarray(tau).reshape(-1, 1)
         theta = 1.0 / tau
         return theta * np.exp(-t * theta) * (t >= 0)
 
     def _integrate(self, t, tau, *args, **kwargs):
+        t = np.asarray(t).reshape(1, -1)
+        tau = np.asarray(tau).reshape(-1, 1)
         theta = 1.0 / tau
         return 1 - np.exp(-t * theta) * (t >= 0)
 
@@ -66,15 +76,20 @@ class MaxwellKernel(UniModalInfluenceKernel):
     That seems not to be autograd differentiable.
     """
     def _call(self, t, tau, *args, **kwargs):
+        t = np.asarray(t).reshape(1, -1)
+        tau = np.asarray(tau).reshape(-1, 1)
         t2 = np.square(t)
         return np.sqrt(2.0/np.pi) * t2 * np.exp(
             -t2 / (2 * tau**2)
         )/(tau**3)
 
     def _mode(self, tau, *args, **kwargs):
+        tau = np.asarray(tau).reshape(-1, 1)
         return np.sqrt(2) * tau
 
[…]

### 2017-03-06 branching_process/cts/influence.py
more generic kernel structure

@@ -17,11 +17,9 @@ class InfluenceKernel(object):
         super(InfluenceKernel, self).__init__(*args)
 
     def majorant(self, t, *args, **kwargs):
-        return np.squeeze(
-            getattr(
+        return getattr(
                 self, '_majorant', self._call
             )(t, *args, **kwargs, **self._fixed_args)
-        )
 
     def __call__(self, t, *args, **kwargs):
         return np.squeeze(
@@ -32,26 +30,6 @@ class InfluenceKernel(object):
         # some kind of numerical quadrature fallback?
         return np.squeeze(
             self._integrate(t, *args, **kwargs, **self._fixed_args)
-        )
-
-
-class UniModalInfluenceKernel(InfluenceKernel):
-    def mode(self, *args, **kwargs):
-        return self._mode(**kwargs, **self._fixed_args)
-
-    def _mode(self, *args, **kwargs):
-        return 0
-
-    def _majorant(self, t, *args, **kwargs):
-        mode = self.mode(*args, **kwargs)
-        peak = self(mode, *args, **kwargs)
-        print('umk', mode, peak)
-        return np.choose(
-            t > mode,
-            [
-                peak,
-                self(t, *args, **kwargs)
-            ]
         )
 
 
@@ -96,67 +74,17 @@ class MaxwellKernel(UniModalInfluenceKernel):
             -np.square(t)/(2 * np.square(tau))
         )
 
-
-class MultiKernel(InfluenceKernel):
-    def __init__(
-            self,
-            n_kernels=1,
-            kernel=MaxwellKernel(),
-            *args, **fixed_args):
-        self.kernel = kernel
-        super(MultiKernel, self).__init__(*args, **fixed_args)
-
-    def majorant(
-            self,
-            t,
-            *args, **kwargs):
-        return np.sum(
-            self.majorant_each(t, *args, **kwargs),
-            0
+    def _majorant(self, t, *args, **kwargs):
+        mode = self.mode(*args, **kwargs)
+        peak = self(mode, *args, **kwargs)
+        print('umk', mode, peak)
+        return np.choose(
+            t > mode,
+            [
+                peak,
+                self(t, *args, **kwargs)
+            ]
         )
-
-    def __call__(
-            self,
-            t,
-            *args, **kwargs):
-        return np.sum(
-            self.call_each(t, *args, **kwargs),
-            0
-        )
-
-    def integrate(
-            self,
-            t,
-            *args, **kwargs):
-        return np.sum(
-            self.integrate_each(t, *args, **kwargs),
[…]

### 2017-03-07 branching_process/cts/influence.py
simplify inheritance

@@ -11,30 +11,79 @@ except ImportError as e:
 
 
 class InfluenceKernel(object):
-
-    def __init__(self, *args, **fixed_args):
+    def __init__(
+            self,
+            n_kernels=1,
+            *args, **fixed_args):
         self._fixed_args = fixed_args
+        self._fixed_args.setdefault('kappa', np.ones(n_kernels)/n_kernels)
         super(InfluenceKernel, self).__init__(*args)
 
-    def majorant(self, t, *args, **kwargs):
+    def majorant(
+            self,
+            t,
+            *args, **kwargs):
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        return np.sum(
+            self.majorant_each(t, *args, **new_kwargs),
+            0
+        )
+
+    def __call__(
+            self,
+            t,
+            *args, **kwargs):
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        return np.sum(
+            self.call_each(t, *args, **new_kwargs),
+            0
+        )
+
+    def integrate(
+            self,
+            t,
+            *args, **kwargs):
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        return np.sum(
+            self.integrate_each(t, *args, **new_kwargs),
+            0
+        )
+
+    def majorant_each(
+            self,
+            t,
+            kappa,
+            *args, **kwargs):
         return getattr(
-                self, '_majorant', self._call
-            )(t, *args, **kwargs, **self._fixed_args)
+            self, '_majorant', self._kernel
+        )(
+            t, *args, **kwargs
+        ) * kappa.reshape(1, -1)
 
-    def __call__(self, t, *args, **kwargs):
-        return np.squeeze(
-            self._call(t, *args, **kwargs, **self._fixed_args)
-        )
+    def call_each(
+            self,
+            t,
+            kappa,
+            *args, **kwargs):
+        return self._kernel(
+            t, *args, **kwargs
+        ) * kappa.reshape(1, -1)
 
-    def integrate(self, t, *args, **kwargs):
-        # some kind of numerical quadrature fallback?
-        return np.squeeze(
-            self._integrate(t, *args, **kwargs, **self._fixed_args)
-        )
+    def integrate_each(
+            self,
+            t,
+            kappa,
[…]

### 2017-03-07 branching_process/cts/influence.py
simpler influence kernel management

@@ -23,75 +23,88 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        new_kwargs = dict()
-        new_kwargs.update(self._fixed_args, **kwargs)
         return np.sum(
-            self.majorant_each(t, *args, **new_kwargs),
-            0
+            self.majorant_each(t, *args, **kwargs),
+            1
         )
 
     def __call__(
             self,
             t,
             *args, **kwargs):
-        new_kwargs = dict()
-        new_kwargs.update(self._fixed_args, **kwargs)
         return np.sum(
-            self.call_each(t, *args, **new_kwargs),
-            0
+            self.call_each(t, *args, **kwargs),
+            1
         )
 
     def integrate(
             self,
             t,
             *args, **kwargs):
-        new_kwargs = dict()
-        new_kwargs.update(self._fixed_args, **kwargs)
         return np.sum(
-            self.integrate_each(t, *args, **new_kwargs),
-            0
+            self.integrate_each(t, *args, **kwargs),
+            1
         )
 
     def majorant_each(
             self,
             t,
-            kappa,
             *args, **kwargs):
+        t = np.asarray(t)
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        tau = np.asarray(new_kwargs.pop('tau'))
+        kappa = np.asarray(new_kwargs.pop('kappa'))
         return getattr(
             self, '_majorant', self._kernel
         )(
-            t, *args, **kwargs
+            t=t.reshape(-1, 1),
+            tau=tau.reshape(1, -1),
+            *args, **new_kwargs
         ) * kappa.reshape(1, -1)
 
     def call_each(
             self,
             t,
-            kappa,
             *args, **kwargs):
+        t = np.asarray(t)
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        tau = np.asarray(new_kwargs.pop('tau'))
+        kappa = np.asarray(new_kwargs.pop('kappa'))
+
         return self._kernel(
-            t, *args, **kwargs
+            t=t.reshape(-1, 1),
+            tau=tau.reshape(1, -1),
+            *args, **new_kwargs
         ) * kappa.reshape(1, -1)
 
[…]

### 2017-03-07 branching_process/cts/influence.py
make sim method compatible with kernels

@@ -137,7 +137,6 @@ class MaxwellKernel(InfluenceKernel):
         t = np.asarray(t)
         mode = np.sqrt(2) * tau
         peak = self._kernel(mode, tau=tau, *args, **kwargs)
-        print('umk', mode, peak)
         return np.choose(
             t > mode,
             [
@@ -163,3 +162,18 @@ class GenericKernel(InfluenceKernel):
         self._majorant = majorant if majorant is not None else kernel
         self._integral = integral
         super(GenericKernel, self).__init__(*args, **fixed_args)
+
+
+def as_influence_kernel(
+        function,
+        majorant=None,
+        integral=None
+    ):
+    if hasattr(function, 'majorant'):
+        return function
+    else:
+        return GenericKernel(
+            kernel=function,
+            majorant=majorant,
+            integral=integral
+        )

### 2017-03-07 branching_process/cts/model.py

@@ -40,7 +40,7 @@ def intensity_hawkes(
             **kwargs)
     deltas = eval_timestamps.reshape(1, -1) - timestamps.reshape(-1, 1)
     mask = deltas > 0.0
-    endo = phi(deltas) * mask
+    endo = phi(deltas.ravel()).reshape(deltas.shape) * mask
     lambdas = endo.sum(0) * eta + mu
     return lambdas
 

### 2017-03-07 branching_process/cts/sim.py

@@ -1,6 +1,7 @@
 import numpy as np
 from numpy import random
 from warnings import warn
+from .  import influence
 
 
 def sim_poisson(
@@ -184,8 +185,7 @@ def sim_branching(
     :rtype: numpy.array
     """
 
-    if phi_m is None:
-        phi_m = getattr(phi, 'majorant', phi)
+    phi = influence.as_influence_kernel(phi, majorant=phi_m)
     Tnext = np.asfarray(immigrants)
     allT = [np.array([])]
 
@@ -194,7 +194,7 @@ def sim_branching(
             lam=phi,
             eta=eta,
             immigrants=Tnext, end=end,
-            lam_m=phi_m)
+            lam_m=phi.majorant)
         # print('gen', gen, Tnext.size)
         if Tnext.size == 0:
             break

### 2017-03-07 branching_process/cts/fit.py
consolidate model fns

@@ -8,7 +8,6 @@ Poisson point process penalised likelihood regression.
 """
 
 try:
-    import autograd
     import autograd.numpy as np
     import autograd.scipy as sp
     have_autograd = True
@@ -16,6 +15,63 @@ except ImportError as e:
     import numpy as np
     import scipy as sp
     have_autograd = False
+
+
+def intensity_hawkes(
+        timestamps,
+        mu,
+        phi,
+        eta=1.0,
+        eval_timestamps=None,
+        sort=True,
+        max_floats=1e8,
+        **kwargs):
+    """
+    True intensity of Hawkes process.
+    Memory-hungry per default; could be improved, with numba.
+    """
+    timestamps = np.asarray(timestamps).ravel()
+    if sort:
+        timestamps = np.sort(timestamps)
+    if eval_timestamps is None:
+        eval_timestamps = timestamps
+        if sort:
+            eval_timestamps = np.sort(eval_timestamps)
+    eval_timestamps = np.asarray(eval_timestamps).ravel()
+    if ((timestamps.size) * (eval_timestamps.size)) > max_floats:
+        return _intensity_hawkes_lite(
+            timestamps=timestamps,
+            mu=mu,
+            phi=phi,
+            eta=eta,
+            eval_timestamps=eval_timestamps,
+            **kwargs)
+    deltas = eval_timestamps.reshape(1, -1) - timestamps.reshape(-1, 1)
+    mask = deltas > 0.0
+    endo = phi(deltas.ravel()).reshape(deltas.shape) * mask
+    lambdas = endo.sum(0) * eta + mu
+    return lambdas
+
+
+def _intensity_hawkes_lite(
+        timestamps,
+        eval_timestamps,
+        mu,
+        phi,
+        eta=1.0,
+        **kwargs):
+    """
+    True intensity of hawkes process.
+    Memory-lite version. CPU-hungry, could be improved with numba.
+    """
+    endo = np.zeros_like(eval_timestamps)
+    deltas = np.zeros_like(timestamps)
+    mask = np.zeros_like(timestamps)
+    for i in range(eval_timestamps.size):
+        deltas[:] = eval_timestamps[i] - timestamps
+        mask[:] = deltas > 0.0
+        endo[i] = np.sum(phi(deltas) * mask)
+    return endo * eta + mu
 
 
 class ContinuousExact(object):

### 2017-03-07 branching_process/cts/model.py

@@ -1,66 +1,0 @@
-try:
-    import autograd.numpy as np
-    import autograd.scipy as sp
-    have_autograd = True
-except ImportError as e:
-    import numpy as np
-    import scipy as sp
-    have_autograd = False
-
-
-
-def intensity_hawkes(
-        timestamps,
-        mu,
-        phi,
-        eta=1.0,
-        eval_timestamps=None,
-        sort=True,
-        max_floats=1e8,
-        **kwargs):
-    """
-    True intensity of Hawkes process.
-    Memory-hungry; could be improved, with numba.
-    """
-    timestamps = np.asarray(timestamps).ravel()
-    if sort:
-        timestamps = np.sort(timestamps)
-    if eval_timestamps is None:
-        eval_timestamps = timestamps
-        if sort:
-            eval_timestamps = np.sort(eval_timestamps)
-    eval_timestamps = np.asarray(eval_timestamps).ravel()
-    if ((timestamps.size) * (eval_timestamps.size)) > max_floats:
-        return _intensity_hawkes_lite(
-            timestamps=timestamps,
-            mu=mu,
-            phi=phi,
-            eta=eta,
-            eval_timestamps=eval_timestamps,
-            **kwargs)
-    deltas = eval_timestamps.reshape(1, -1) - timestamps.reshape(-1, 1)
-    mask = deltas > 0.0
-    endo = phi(deltas.ravel()).reshape(deltas.shape) * mask
-    lambdas = endo.sum(0) * eta + mu
-    return lambdas
-
-
-def _intensity_hawkes_lite(
-        timestamps,
-        eval_timestamps,
-        mu,
-        phi,
-        eta=1.0,
-        **kwargs):
-    """
-    True intensity of hawkes process.
-    Memory-lite version. CPU-hungry, could be improved with numba.
-    """
-    endo = np.zeros_like(eval_timestamps)
-    deltas = np.zeros_like(timestamps)
-    mask = np.zeros_like(timestamps)
-    for i in range(eval_timestamps.size):
-        deltas[:] = eval_timestamps[i] - timestamps
-        mask[:] = deltas > 0.0
-        endo[i] = np.sum(phi(deltas) * mask)
-    return endo * eta + mu

### 2017-03-08 branching_process/cts/influence.py
can haz StepKernel

@@ -13,10 +13,10 @@ except ImportError as e:
 class InfluenceKernel(object):
     def __init__(
             self,
-            n_kernels=1,
+            n_bases=1,
             *args, **fixed_args):
         self._fixed_args = fixed_args
-        self._fixed_args.setdefault('kappa', np.ones(n_kernels)/n_kernels)
+        self._fixed_args.setdefault('kappa', np.ones(n_bases)/n_bases)
         super(InfluenceKernel, self).__init__(*args)
 
     def majorant(
@@ -144,6 +144,71 @@ class MaxwellKernel(InfluenceKernel):
                 self._kernel(t, tau=tau, *args, **kwargs)
             ]
         )
+
+
+class StepKernel(InfluenceKernel):
+    """
+    Piecewise-constant rate.
+    This is presumed to be for background rate modelling.
+    """
+    def __init__(
+            self,
+            end,
+            n_bases=100,
+            *args,
+            **fixed_args
+            ):
+        self.end = end
+        super(StepKernel, self).__init__(n_bases=n_bases, *args, **fixed_args)
+        self._fixed_args.setdefault(
+            'tau',
+            np.linspace(0, end, n_bases+1, endpoint=True)
+        )
+
+    def __call__(self, t, *args, **kwargs):
+        """
+        because we need n_bases+1 tau points for this, it doesn't map onto
+        the easy structure of the others.
+        I'm not convinced this subclassing rigmarole is worth the effort.
+        """
+        t = np.asarray(t)
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        tau = np.asarray(new_kwargs.pop('tau'))
+        mu = new_kwargs.pop('mu', 0.0)
+        kappa = np.asarray(new_kwargs.pop('kappa'))
+        kappa = np.maximum(kappa, -mu)
+        t = t.reshape(-1, 1)
+        each = (
+            (t > tau[:-1].reshape(1, -1)) -
+            (t > tau[1:].reshape(1, -1))
+        )
+        return np.sum(
+            each * kappa.reshape(1, -1),
+            1
+        ) + mu
+
+    def integrate(self, t, *args, **kwargs):
+        t = np.asarray(t)
+        new_kwargs = dict()
+        new_kwargs.update(self._fixed_args, **kwargs)
+        tau = np.asarray(new_kwargs.pop('tau'))
[…]

### 2017-03-08 branching_process/cts/model.py
less glm-structured thing

@@ -16,6 +16,7 @@ except ImportError as e:
     import scipy as sp
     have_autograd = False
 
+from . import influence
 
 def intensity_hawkes(
         timestamps,
@@ -77,15 +78,17 @@ def _intensity_hawkes_lite(
 class ContinuousExact(object):
     def __init__(
             self,
-            phis,
+            phi=None,
             fit_tau=False
             ):
-        if hasattr(phis, 'integrate'):
-            # assume a single kernel, instead of an array
-            phis = [phis]
-        self.phis = phis
-        self.n_bases = len(phis)
+        if phi is None:
+            phi = influence.MaxwellKernel(n_bases=5)
+        else:
+            phi = influence.as_influence_kernel(phi)
+        self.phi = phi
+        self.n_bases = phi.n_bases
         self.fit_tau = fit_tau
+        # we always fit kappa
 
     def _pack(
             self,
@@ -134,6 +137,7 @@ class ContinuousExact(object):
             taus=0.0,
             log_omegas=0.0,
             ):
+        
         endo_rate = np.dot(np.reshape(kappas, (1, -1)), X)
         lamb = endo_rate + mu * np.exp(log_omega)
         partial_loglik = loglik_poisson(lamb, y)
@@ -172,7 +176,8 @@ class ContinuousExact(object):
             # pi_omega=0.0,
             ):
         """
-        approximate self._dof_packed differentiably, using gradient information.
+        approximate self._dof_packed differentiably,
+        using gradient information.
         """
         raise NotImplementedError()
 
@@ -227,11 +232,11 @@ class ContinuousExact(object):
             t0=0.0,
             t1=None,
             pi_kappa=0.0,
-            # pi_omega=1e-8,
-            # max_steps=None,
-            # step_iter=50,
-            # step_size=0.1,
-            # gamma=0.9,
+            pi_omega=1e-8,
+            max_steps=None,
+            step_iter=50,
+            step_size=0.1,
+            gamma=0.9,
             eps=1e-8,
             # backoff=0.75
             ):
@@ -239,7 +244,13 @@ class ContinuousExact(object):
             param_vector = self._pack()
 
         self.t0 = t0
-        self.t1 = t1 or np.maximum(ts)
+        self.t1 = t1 or ts[-1]
+
+        # Full likelihood is best evaluated at the end also
+        if ts[-1] < self.t1:
+            evalpts = np.append(ts[ts>t0], [self.t1])
+        else:
[…]

### 2017-03-08 branching_process/cts/influence.py
nearly coherent

@@ -17,6 +17,7 @@ class InfluenceKernel(object):
             *args, **fixed_args):
         self._fixed_args = fixed_args
         self._fixed_args.setdefault('kappa', np.ones(n_bases)/n_bases)
+        self.n_bases = n_bases
         super(InfluenceKernel, self).__init__(*args)
 
     def majorant(
@@ -146,7 +147,7 @@ class MaxwellKernel(InfluenceKernel):
         )
 
 
-class StepKernel(InfluenceKernel):
+class StepLinearKernel(InfluenceKernel):
     """
     Piecewise-constant rate.
     This is presumed to be for background rate modelling.
@@ -159,7 +160,7 @@ class StepKernel(InfluenceKernel):
             **fixed_args
             ):
         self.end = end
-        super(StepKernel, self).__init__(n_bases=n_bases, *args, **fixed_args)
+        super(StepLinearKernel, self).__init__(n_bases=n_bases, *args, **fixed_args)
         self._fixed_args.setdefault(
             'tau',
             np.linspace(0, end, n_bases+1, endpoint=True)
@@ -232,13 +233,15 @@ class GenericKernel(InfluenceKernel):
 def as_influence_kernel(
         function,
         majorant=None,
-        integral=None
-    ):
+        integral=None,
+        n_bases=1
+        ):
     if hasattr(function, 'majorant'):
         return function
     else:
         return GenericKernel(
             kernel=function,
             majorant=majorant,
-            integral=integral
+            integral=integral,
+            n_bases=n_bases
         )

### 2017-03-08 branching_process/cts/model.py

@@ -8,6 +8,7 @@ Poisson point process penalised likelihood regression.
 """
 
 try:
+    import autograd
     import autograd.numpy as np
     import autograd.scipy as sp
     have_autograd = True
@@ -79,16 +80,14 @@ class ContinuousExact(object):
     def __init__(
             self,
             phi=None,
-            fit_tau=False
+            n_bases=1
             ):
         if phi is None:
             phi = influence.MaxwellKernel(n_bases=5)
         else:
-            phi = influence.as_influence_kernel(phi)
+            phi = influence.as_influence_kernel(phi, n_bases=n_bases)
         self.phi = phi
         self.n_bases = phi.n_bases
-        self.fit_tau = fit_tau
-        # we always fit kappa
 
     def _pack(
             self,
@@ -98,11 +97,11 @@ class ContinuousExact(object):
             # log_omega=0.0,
             ):
 
-        n_taus = self.n_bases * self.fit_tau
+        n_taus = self.n_bases * self._fit_tau
         packed = np.zeros(
             1 +
             self.n_bases +
-            self.n_bases * self.fit_tau
+            self.n_bases * self._fit_tau
             # n_steps
         )
         packed[0] = mu
@@ -117,7 +116,7 @@ class ContinuousExact(object):
         """
         returns mu, kappas, taus, #log_omegas
         """
-        n_taus = self.n_bases * self.fit_tau
+        n_taus = self.n_bases * self._fit_tau
         return (
             packed[0],
             packed[1:self.n_bases+1],
@@ -128,16 +127,25 @@ class ContinuousExact(object):
     def _negloglik_packed(
             self,
             param_vector):
-        return self.negloglik(*self._unpack(param_vector))
+        return self.negloglik(
+            self._ts, self._evalpts, phi=self.phi,
+            *self._unpack(param_vector))
 
     def negloglik(
             self,
-            mu,
-            kappas,
+            ts,
+            phi=None,
+            evalpts=None,
+            mu=1.0,
+            kappas=None,
             taus=0.0,
             log_omegas=0.0,
             ):
-        
+        if phi is None:
+            phi = self.phi
+        if kappas is None:
+            kappas = np.ones(phi.n_bases) / phi.n_bases
+
         endo_rate = np.dot(np.reshape(kappas, (1, -1)), X)
         lamb = endo_rate + mu * np.exp(log_omega)
         partial_loglik = loglik_poisson(lamb, y)
[…]

### 2017-03-08 branching_process/cts/sim.py

@@ -218,6 +218,7 @@ def sim_hawkes(
         start=0.0, end=1.0,
         phi_m=None,
         immigrants=None,
+        sort=True,
         **kwargs):
     """
     Basic Hawkes process
@@ -263,7 +264,7 @@ def sim_hawkes(
         immigrants,
         sim_poisson(mu, start, end)
     )
-    return np.append(
+    ts = np.append(
         immigrants,
         sim_branching(
             immigrants=immigrants,
@@ -274,3 +275,6 @@ def sim_hawkes(
             **kwargs
         )
     )
+    if sort:
+        ts = np.sort(ts)
+    return ts

### 2017-03-08 branching_process/cts/model.py
rework args

@@ -80,48 +80,51 @@ class ContinuousExact(object):
     def __init__(
             self,
             phi=None,
-            n_bases=1
+            n_phi_bases=1,
+            n_omega_bases=0,
             ):
         if phi is None:
             phi = influence.MaxwellKernel(n_bases=5)
         else:
-            phi = influence.as_influence_kernel(phi, n_bases=n_bases)
+            phi = influence.as_influence_kernel(phi, n_bases=n_phi_bases)
         self.phi = phi
-        self.n_bases = phi.n_bases
+        self.n_phi_bases = phi.n_bases
+        self.n_omega_bases = n_omega_bases
 
     def _pack(
             self,
             mu=1.0,
-            kappas=0.0,
-            taus=[]
-            # log_omega=0.0,
+            kappa=0.0,
+            tau=1.0,
+            log_omega=0.0,
             ):
 
-        n_taus = self.n_bases * self._fit_tau
+        n_tau = self.n_phi_bases * self._fit_tau
+        n_omegas = self.n_omega_bases * self._fit_omega
         packed = np.zeros(
             1 +
-            self.n_bases +
-            self.n_bases * self._fit_tau
-            # n_steps
+            self.n_phi_bases +
+            n_tau +
+            n_omegas
         )
         packed[0] = mu
-        packed[1:self.n_bases + 1] = kappas
-        packed[self.n_bases + 1:self.n_bases + n_taus + 1] = taus
-        # packed[self.n_bases+n_taus + 1:] = log_omega
+        packed[1:self.n_phi_bases + 1] = kappa
+        packed[self.n_phi_bases + 1:self.n_phi_bases + n_tau + 1] = tau
+        packed[self.n_phi_bases + n_tau + 1:] = log_omega
         return packed
 
     def _unpack(
             self,
             packed):
         """
-        returns mu, kappas, taus, #log_omegas
+        returns mu, kappa, tau, #log_omega
         """
-        n_taus = self.n_bases * self._fit_tau
+        n_tau = self.n_phi_bases * self._fit_tau
         return (
             packed[0],
-            packed[1:self.n_bases+1],
-            packed[self.n_bases + 1:self.n_bases + n_taus + 1],
-            # packed[self.n_bases + n_taus + 1:]
+            packed[1:self.n_phi_bases+1],
+            packed[self.n_phi_bases + 1:self.n_phi_bases + n_tau + 1],
+            packed[self.n_phi_bases + n_tau + 1:]
         )
 
     def _negloglik_packed(
[…]

### 2017-03-08 branching_process/cts/model.py
attempt real loglik calculations

@@ -131,14 +131,16 @@ class ContinuousExact(object):
             self,
             param_vector):
         return self.negloglik(
-            self._ts, self._evalpts, phi=self.phi,
+            self._ts,
+            self._evalpts,
+            self.phi,
             *self._unpack(param_vector))
 
     def negloglik(
             self,
             ts,
-            phi=None,
             evalpts=None,
+            phi=None,
             mu=1.0,
             kappa=None,
             tau=0.0,
@@ -148,11 +150,10 @@ class ContinuousExact(object):
             phi = self.phi
         if kappa is None:
             kappa = np.ones(phi.n_omega_bases) / phi.n_omega_bases
-
-        endo_rate = np.dot(np.reshape(kappa, (1, -1)), X)
-        lamb = endo_rate + mu * np.exp(log_omega)
-        partial_loglik = loglik_poisson(lamb, y)
-        return -np.sum(partial_loglik)
+        lam = phi(ts, tau=tau) + mu
+        big_lam = phi.integral(self._t_end) - phi.integral(self._t_start)
+        negloglik = big_lam - np.sum(np.log(lam))
+        return negloglik
 
     def _penalty_weight_packed(
             self,

### 2017-03-08 branching_process/cts/model.py
snapshot of optional optimisation params

@@ -143,14 +143,14 @@ class ContinuousExact(object):
             phi=None,
             mu=1.0,
             kappa=None,
-            tau=0.0,
-            log_omega=0.0,
+            tau=[],
+            log_omega=[],
             ):
         if phi is None:
             phi = self.phi
         if kappa is None:
             kappa = np.ones(phi.n_omega_bases) / phi.n_omega_bases
-        lam = phi(ts, tau=tau) + mu
+        lam = phi(ts, tau=tau, kappa=kappa) + mu
         big_lam = phi.integral(self._t_end) - phi.integral(self._t_start)
         negloglik = big_lam - np.sum(np.log(lam))
         return negloglik

### 2017-03-08 branching_process/cts/influence.py
massage arguments better

@@ -51,11 +51,11 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asarray(t)
+        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asarray(new_kwargs.pop('tau'))
-        kappa = np.asarray(new_kwargs.pop('kappa'))
+        tau = np.asfarray(new_kwargs.pop('tau'))
+        kappa = np.asfarray(new_kwargs.pop('kappa'))
         return getattr(
             self, '_majorant', self._kernel
         )(
@@ -68,11 +68,11 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asarray(t)
+        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asarray(new_kwargs.pop('tau'))
-        kappa = np.asarray(new_kwargs.pop('kappa'))
+        tau = np.asfarray(new_kwargs.pop('tau'))
+        kappa = np.asfarray(new_kwargs.pop('kappa'))
 
         return self._kernel(
             t=t.reshape(-1, 1),
@@ -84,11 +84,11 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asarray(t)
+        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asarray(new_kwargs.pop('tau'))
-        kappa = np.asarray(new_kwargs.pop('kappa'))
+        tau = np.asfarray(new_kwargs.pop('tau'))
+        kappa = np.asfarray(new_kwargs.pop('kappa'))
         return self._integrate(
             t=t.reshape(-1, 1),
             tau=tau.reshape(1, -1),
@@ -98,14 +98,14 @@ class InfluenceKernel(object):
 
 class ExpKernel(InfluenceKernel):
     def _kernel(self, t, tau, *args, **kwargs):
-        t = np.asarray(t)
-        tau = np.asarray(tau)
+        t = np.asfarray(t)
+        tau = np.asfarray(tau)
         theta = 1.0 / tau
         return theta * np.exp(-t * theta) * (t >= 0)
 
[…]

### 2017-03-08 branching_process/cts/model.py

@@ -32,14 +32,14 @@ def intensity_hawkes(
     True intensity of Hawkes process.
     Memory-hungry per default; could be improved, with numba.
     """
-    timestamps = np.asarray(timestamps).ravel()
+    timestamps = np.asfarray(timestamps).ravel()
     if sort:
         timestamps = np.sort(timestamps)
     if eval_timestamps is None:
         eval_timestamps = timestamps
         if sort:
             eval_timestamps = np.sort(eval_timestamps)
-    eval_timestamps = np.asarray(eval_timestamps).ravel()
+    eval_timestamps = np.asfarray(eval_timestamps).ravel()
     if ((timestamps.size) * (eval_timestamps.size)) > max_floats:
         return _intensity_hawkes_lite(
             timestamps=timestamps,
@@ -100,12 +100,12 @@ class ContinuousExact(object):
             ):
 
         n_tau = self.n_phi_bases * self._fit_tau
-        n_omegas = self.n_omega_bases * self._fit_omega
+        n_omega = self.n_omega_bases * self._fit_omega
         packed = np.zeros(
             1 +
             self.n_phi_bases +
             n_tau +
-            n_omegas
+            n_omega
         )
         packed[0] = mu
         packed[1:self.n_phi_bases + 1] = kappa
@@ -120,12 +120,20 @@ class ContinuousExact(object):
         returns mu, kappa, tau, #log_omega
         """
         n_tau = self.n_phi_bases * self._fit_tau
-        return (
-            packed[0],
-            packed[1:self.n_phi_bases+1],
-            packed[self.n_phi_bases + 1:self.n_phi_bases + n_tau + 1],
-            packed[self.n_phi_bases + n_tau + 1:]
-        )
+        n_omega = self.n_omega_bases * self._fit_omega
+        unpacked = {
+            'mu': packed[0],
+            'kappa': packed[1:self.n_phi_bases+1],
+        }
+        if n_tau > 0:
+            unpacked['tau'] = packed[
+                self.n_phi_bases + 1:self.n_phi_bases + n_tau + 1
+            ]
+        if n_omega > 0:
+            unpacked['omega'] = packed[
+                self.n_phi_bases + n_tau + 1:
+            ]
+        return unpacked
 
     def _negloglik_packed(
             self,
@@ -134,7 +142,7 @@ class ContinuousExact(object):
             self._ts,
             self._evalpts,
             self.phi,
-            *self._unpack(param_vector))
[…]

### 2017-03-08 branching_process/cts/influence.py
reshape through numpy api

@@ -59,10 +59,10 @@ class InfluenceKernel(object):
         return getattr(
             self, '_majorant', self._kernel
         )(
-            t=t.reshape(-1, 1),
-            tau=tau.reshape(1, -1),
+            t=np.reshape(t, (-1, 1)),
+            tau=np.reshape(tau, (1, -1)),
             *args, **new_kwargs
-        ) * kappa.reshape(1, -1)
+        ) * np.reshape(kappa, (1, -1))
 
     def call_each(
             self,
@@ -75,10 +75,10 @@ class InfluenceKernel(object):
         kappa = np.asfarray(new_kwargs.pop('kappa'))
 
         return self._kernel(
-            t=t.reshape(-1, 1),
-            tau=tau.reshape(1, -1),
+            t=np.reshape(t, (-1, 1)),
+            tau=np.reshape(tau, (1, -1)),
             *args, **new_kwargs
-        ) * kappa.reshape(1, -1)
+        ) * np.reshape(kappa, (1, -1))
 
     def integrate_each(
             self,
@@ -90,10 +90,10 @@ class InfluenceKernel(object):
         tau = np.asfarray(new_kwargs.pop('tau'))
         kappa = np.asfarray(new_kwargs.pop('kappa'))
         return self._integrate(
-            t=t.reshape(-1, 1),
-            tau=tau.reshape(1, -1),
+            t=np.reshape(t, (-1, 1)),
+            tau=np.reshape(tau, (1, -1)),
             *args, **new_kwargs
-        ) * kappa.reshape(1, -1)
+        ) * np.reshape(kappa, (1, -1))
 
 
 class ExpKernel(InfluenceKernel):
@@ -179,13 +179,13 @@ class StepLinearKernel(InfluenceKernel):
         mu = new_kwargs.pop('mu', 0.0)
         kappa = np.asfarray(new_kwargs.pop('kappa'))
         kappa = np.maximum(kappa, -mu)
-        t = t.reshape(-1, 1)
+        t = np.reshape(t, (-1, 1))
         each = (
             (t > tau[:-1].reshape(1, -1)) -
             (t > tau[1:].reshape(1, -1))
         )
         return np.sum(
-            each * kappa.reshape(1, -1),
+            each * np.reshape(kappa, (1, -1)),
             1
         ) + mu
 
@@ -197,7 +197,7 @@ class StepLinearKernel(InfluenceKernel):
         mu = new_kwargs.pop('mu', 0.0)
[…]

### 2017-03-08 branching_process/cts/influence.py
no more np.asfarray
array casting confuses autograd

@@ -51,11 +51,10 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asfarray(new_kwargs.pop('tau'))
-        kappa = np.asfarray(new_kwargs.pop('kappa'))
+        tau = new_kwargs.pop('tau')
+        kappa = new_kwargs.pop('kappa')
         return getattr(
             self, '_majorant', self._kernel
         )(
@@ -68,11 +67,10 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asfarray(new_kwargs.pop('tau'))
-        kappa = np.asfarray(new_kwargs.pop('kappa'))
+        tau = new_kwargs.pop('tau')
+        kappa = new_kwargs.pop('kappa')
 
         return self._kernel(
             t=np.reshape(t, (-1, 1)),
@@ -84,11 +82,10 @@ class InfluenceKernel(object):
             self,
             t,
             *args, **kwargs):
-        t = np.asfarray(t)
         new_kwargs = dict()
         new_kwargs.update(self._fixed_args, **kwargs)
-        tau = np.asfarray(new_kwargs.pop('tau'))
-        kappa = np.asfarray(new_kwargs.pop('kappa'))
+        tau = new_kwargs.pop('tau')
+        kappa = new_kwargs.pop('kappa')
         return self._integrate(
             t=np.reshape(t, (-1, 1)),
             tau=np.reshape(tau, (1, -1)),
@@ -98,14 +95,10 @@ class InfluenceKernel(object):
 
 class ExpKernel(InfluenceKernel):
     def _kernel(self, t, tau, *args, **kwargs):
-        t = np.asfarray(t)
-        tau = np.asfarray(tau)
         theta = 1.0 / tau
         return theta * np.exp(-t * theta) * (t >= 0)
 
     def _integrate(self, t, tau, *args, **kwargs):
-        t = np.asfarray(t)
-        tau = np.asfarray(tau)
         theta = 1.0 / tau
         return 1 - np.exp(-t * theta) * (t >= 0)
 
[…]

### 2017-03-08 branching_process/cts/__init__.py
now, does this thing actually give results?

@@ -1,3 +1,3 @@
 from __future__ import absolute_import
 
-from .fit import ContinuousExact
+from .model import ContinuousExact

### 2017-03-08 branching_process/cts/model.py

@@ -82,6 +82,7 @@ class ContinuousExact(object):
             phi=None,
             n_phi_bases=1,
             n_omega_bases=0,
+            debug=False
             ):
         if phi is None:
             phi = influence.MaxwellKernel(n_bases=5)
@@ -90,15 +91,20 @@ class ContinuousExact(object):
         self.phi = phi
         self.n_phi_bases = phi.n_bases
         self.n_omega_bases = n_omega_bases
+        self.debug = debug
+
+    def _debug_print(self, *args, **kwargs):
+        if self.debug:
+            print(*args, **kwargs)
 
     def _pack(
             self,
             mu=1.0,
-            kappa=0.0,
+            kappa=0.1,
             tau=1.0,
             log_omega=0.0,
+            **kwargs
             ):
-
         n_tau = self.n_phi_bases * self._fit_tau
         n_omega = self.n_omega_bases * self._fit_omega
         packed = np.zeros(
@@ -137,12 +143,13 @@ class ContinuousExact(object):
 
     def _negloglik_packed(
             self,
-            param_vector):
+            param_vector, **kwargs):
         return self.negloglik(
             self._ts,
             self._evalpts,
             self.phi,
-            **self._unpack(param_vector))
+            **self._unpack(param_vector),
+            **kwargs)
 
     def negloglik(
             self,
@@ -186,7 +193,7 @@ class ContinuousExact(object):
             np.abs(param_vector) *
             self._penalty_weight_packed(
                 pi_kappa,
-                # pi_omega
+                pi_omega
             )
         )
 
@@ -229,10 +236,6 @@ class ContinuousExact(object):
     def fit(
             self,
             ts,
-            mu=0.0,
-            kappa=0.0,
-            tau=0.0,
-            log_omega=0.0,
             fit_tau=False,
             fit_omega=False,
             **kwargs
@@ -242,10 +245,7 @@ class ContinuousExact(object):
         return self._fit_packed(
             ts,
             self._pack(
-                mu=mu,
-                kappa=kappa,
-                tau=tau,
-                log_omega=log_omega
+                **kwargs
             ),
             **kwargs
         )
@@ -295,22 +295,22 @@ class ContinuousExact(object):
         self._evalpts = _evalpts
 
         param_floor = self._pack(
-            mu=0.0,  # unused
+            mu=0.0,  # unused?
             kappa=0.0,
             log_omega=-np.inf
         )
 
[…]

### 2017-03-08 branching_process/cts/model.py
my negloglik function is not plausible

@@ -19,6 +19,7 @@ except ImportError as e:
 
 from . import influence
 
+
 def intensity_hawkes(
         timestamps,
         mu,
@@ -157,19 +158,31 @@ class ContinuousExact(object):
             evalpts=None,
             phi=None,
             mu=1.0,
+            start=None,
+            end=None,
             kappa=None,
             log_omega=[],
             **kwargs):
         if phi is None:
             phi = self.phi
+        if end is None:
+            end = getattr(self, '_t_end', ts[-1])
+        if start is None:
+            start = getattr(self, '_t_start', 0.0)
         lam = phi(ts, kappa=kappa, **kwargs) + mu
         big_lam = phi.integrate(
-            self._t_end, kappa=kappa, **kwargs
+            end, kappa=kappa, **kwargs
             ) - phi.integrate(
-            self._t_start, kappa=kappa, **kwargs
+            start, kappa=kappa, **kwargs
         )
         negloglik = big_lam - np.sum(np.log(lam))
         return negloglik
+
+    def loglik(
+            self,
+            *args,
+            **kwargs):
+        return -self.loglik(*args, **kwargs)
 
     def _penalty_weight_packed(
             self,

### 2017-04-24 _paper_output/paper_sparse_hawkes.tex
more conveneient paper ignoring

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters

\title{Regularised parameter selection in the non-stationary Hawkes process}

\author{Daniel MacKinlay}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{$\ell_1$-penalised non-stationary Hawkes branching
process}\label{ellux5f1-penalised-non-stationary-hawkes-branching-process}

    \subsection{Abstract}\label{abstract}

We are interested in the problem of simultaneous identification of the
parameters of a Hawkes point process and identifying a sparse driving
signal by means of $\ell_1$ penalised regression where, by assumption,
the driving noise has a different form. This is analaogous to a blind
deconvolution problem or a compressive sensing-style inverse problem,
but considerably more challenging; While the process model superfcially
resembles a linear system, there are few of the conveneient results that
woudl apply in a linear regression problem; and as a time series problem
it is difficult to choose our penalty parameters by standard
non-parametric means such as cross validation. The proposed method uses
a heuristically-motivated Akaike-style Information criterion
regularisation parameter selection, which we test through simulation.
The proposed method works well in practice.

    \subsection{Introduction}\label{introduction}

We define call a counting process $N$ indexed by $t$ a (TODO)
predictable Poisson point process if it has an associated
\emph{intensity} process $\lambda$, such that $\lambda$ is predictable
with respect to the filtration generated by $N$, i.e.
\[\lambda(\mathcal{F(t^-)}):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]

The Hawkes process (Hawkes, 1971) specifically, is the linear,
self-exciting point process; that is, given a "background" rate
$\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,$ and a non-negative influence
kernel with positive support
$\phi:\mathbb{R}^+\rightarrow\mathbb{R}^+\cap\{0\}$, the Hawkes process
has the following expression for its intensity:

\[
\begin{aligned}
\lambda(\mathcal{F(t)}^-_{(-\infty,t)}) &= \mu(t) + \phi_\theta * N\\
&= \mu(t)  + \int_{-\infty}^{\infty}\phi_\theta(t-s)dN(s)\\
&= \mu(t) + \int_{-\infty}^{t}\phi_\theta(t-s)dN(s)\\
&= \mu(t) + \sum_{t_i\lt t}\phi_\theta(t-t_i)
\end{aligned}
\]

In this problem, we are given a signal realisation of the count process
$N$ on an interval $T=[t_0,t_N]$. We are interested in identifying, from
this data, two unknown functional parameters; both the background rate
function $\mu$ and the influence kernel $\phi$.

    One such assumption would be penalization :citeTikhonov1965,Hoerl1970
-{}-I apply an additional penalty to the log likelihood function,
penalizing particular values of the parameters.

For the penalized log likelihood for parameter $\theta$ and penalty
$\pi$, I write

\[L_\pi(\mathbf t, \theta'):= L_\pi(\mathbf t, \theta')- \pi P(\theta')\]

$P$ here is a non-negative functional which penalizes certain values of
the parameter vector, and $\pi\geq 0$ is the penalty weight
hyperparameter.

As before, the estimate is defined as the maximizer:

\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta)\]

For non-parametric extension to a parametric model, one usually
penalizes only the non-parametric extensions to the model, such that
they vanish as the penalty term increases. :citeGreen1987

\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\mathbf t) =\hat{\theta}(\mathbf t)\]

I follow this practice here, penalizing only the deviations of
$\omega(t)$ from the parametric estimate.

Hereafter, I will drop the prime from the augmented parameter vector and
simply write $\theta$.

Penalization is more frequently presented in the context of generalized
linear regression from i.id. samples, but it also fits within a
generalized Maximum Likelihood estimation theory.

Various gradient ascent algorithms are known to perform well for
:cite::Simon2011,Wu2008. Pragmatically, for performance reasons, I used
a mixed strategy. My algorithm attempts to update marginal estimates for
each $\omega_i$ parameter more rapidly first via Newton's method
:citeBattiti1992,Ozaki1979 and uses conjugate gradient descent for the
parametric estimates of the Hawkes parameters. If the updates steps are
small this seems to be stable. There are various tuning parameters to
such algorithms.

Many variants of pathwise regularization and exist, such as backwards
stepwise regularization, versions with random initialization, Least
Angle Regression, :citeEfron2004a selection with integrated cross
validation and so on. :citeFriedman2010

Many variants of pathwise regularization and exist, such as backwards
stepwise regularization, versions with random initialization, Least
Angle Regression, :citeEfron2004a selection with integrated cross
validation and so on. :citeFriedman2010 For this particular model I have
found few results giving me analytic guarantees, so I use heuristics and
simulation-based verification. I do not declare that this method "best"
in any sense, or that it will find the correct global maximum penalized
likelihood etc. The method, however, is simple enough to prototype
rapidly and, as it turns out, performs well on test data. Results will
be presented in the next chapter.

    We take $\mu$ to be an unknown sparse, non-random signal in a particular
sense. We will assume that, given a linear basis
$\{\psi_j \}_{1\leq j\leq M}$, that we may write the background rate as
\[\mu(t) = \mu_0 + \sum_{j=1}^M  \omega_j \psi_j(t).\] \emph{Sparsity}
here will imply that the coefficients $\omega_j$ will be mostly 0.

In practice, we choose
\[\psi_j(t) := \mathbb{I}_{[\tau_{j},\tau_{j+1})}(t)\] for some lattice
$\{\tau_j\},$ which will correspond to the assumption that "sparse"
triggers will have localised support in time; we are looking for short
"spikes".

Similarly, we will approximate $\phi$ in a different functional basis
${f}_{1\leq k \leq K}$, such that $\phi(s):= \sum_{k=1}^K f(s;k)$. This
choice is likely arbitrary, but many exist in the literature trading off
computational and statistical convenience, including Laguerre
polynomials (Ogata and Akaike, 1982), Exponential functions (Schoenberg
2015; Rambaldi et al, 2015), and step functions (Eichler, 2016). As a
convenient choice, we propose a basis of smooth unimodal non-negative
functions with a computationally convenient form and a single mode, the
convenient Maxwell distribution,
\[\phi_j(t):=\sqrt{\frac{2}{\pi}} \frac{t^2\exp\left(-\frac{t^2}{2\sigma_j^2}\right)}{\sigma_j^3}.\]

A motivation for these specific choices is a particular real-world
problem: modelling viral memes in social media. We suspect that online
popularity is governed by some mixture of organic social sharing,
modeled by the self-exciting $\phi$ component, mixed with intermittent
external boosts, such as when an internet meme is featured on an
external medium such as television.

    mean delay $\mu=2\tau \sqrt{\frac{2}{\pi}}$

mode $\sqrt{2}\tau$

scale $\sigma=\sqrt{\frac{\tau^2(3 \pi - 8)}{\pi}}$

We take fixed $M$-point lattice $\tilde{\tau}$ and $K$-fixed scales
$\tilde{\sigma}.$

The parameter vector to be estimated then is
$\theta:=(\mu, \tilde{\omega}, \tilde{\kappa}).$

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_10_0.pdf}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Optimization}\label{optimization}

The non-penalised log-likelihood function for a given realisation of the
data is given

\[L_\pi(\mathbf t;\theta):=-\int_0^T\lambda_\theta(t)dt + \int_0^T\log \lambda_\theta(t) dN_t - \pi\|\boldsymbol \omega_\theta(t)\|_1\]

And using the method of maximum likelihood we estimate our parameter
vector by numerical optimisation of
\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta).\]

There is no closed-form solution for the optimum, but the derivatives of
this function are easy to calculate manually see e.g. (Ozaki 1979). Here
howevfer, we use automatic differentiation to make more complex
penalisation procedure conveneient to differentiate (Baydin and
Pearlmutter 2014).

We are able to use these to devise rapid gradient-pased optimisation
procedure.

    To remedy this, I use Sugiura's finite-sample-corrected version, the
\emph{AICc} (Sugiura, 1978).
\[{\mathrm {AIC}(L(X;\hat{\theta}))}=2d-2\ln(L)\]

\[{\mathrm {AICc}}:={\mathrm {AIC}}+{\frac {2d(d+1)}{N-d-1}}\]

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_13_0.pdf}
    \end{center}
    { \hspace*{\fill} \\}

    \section{Hyperparameter selection}\label{hyperparameter-selection}

    \subsubsection{Degrees of freedom}\label{degrees-of-freedom}

Whilst we might wish to select the regularisation penalties by cross
validation, cross validation is not straightforward for time series.

The traditional means of performing model selection in Hawkes-type
models e.g. Ogata (1983), uses \emph{degrees-of-freedom} penalties
(Efron 1986). Specifically, the \emph{AIC}, the Akaike Information
Criterion (Akaike 1981) is widely popular.

The AIC formula, for a model $M$ fit to a given data set $X$, for
estimated parameter vector $\hat{\theta^M}$ with log likelihood $L$ and
degrees of freedom $d^M$

\[\mathrm{AIC}(X,M)=2d^M-2L^M(X,\hat{\theta}^M)\]

The degrees of freedom for many models taken to be equal to the
cardinality of the parameter vector $\theta$.

Given two ML fitted models, $(M_1,\hat{\theta}^M_1)$ and
$(M_2\,\hat{\theta}^M_2)$, the difference in their AIC values is an
estimate of the relative Kullback-Leibler divergence of the inferred
measures $\hat{\mu_1},\hat{\mu_2}$ from the unknown true distribution,
$\mu$ i.e.

\[\mathrm{AIC}(X,M_1)-\mathrm{AIC}(X,M_2) \simeq D_\mathrm{KL}(\mu\|\hat{\mu_1}) - D_\mathrm{KL}(\mu\|\hat{\mu_2})\]

However, the AIC does not immediately solve the problem of
regularizition parameter selection in maximum likelihood, where the
regularisation parameter may be a real value smoothly interpolating many
possible models. One can calculate an AIC for penalised models, but this
require onerous Hessian matrix inversions and becomes costly in high
dimensions, and in any case does not generalise readily to non-smooth
penalties. (Konishi and kitagawa 2008)

We proceed heuristically, based on the arguments from Lasso-type
linear-model regularisation. There, the number of non-zero coefficients
fit under a particular value of the regularization penalty $\pi$ in an
$\ell_1$ regression model provides an unbiassed estimate of the
effective degrees of freedom of the model (Zou et al, 2007.)

    This suggests an adjusted likelihood parameter of the following form:

\[
\widehat{\mathrm{AICc}}(X, \hat{\theta}_\pi) = {\frac {2Nd(\hat{\theta}_\pi) }{N-d(\hat{\theta}_\pi)-1}} - 2\ln(L(X;\hat{\theta}_\pi), X))
\]

where $d_\pi$ counts the number of non-zero coefficients estimated for
given regularization coefficient $\pi$.

    Extensive grid search for optimsal parameters is expensive, so we select
both regularisation parameters using optimal sampling by Gaussian
process regression of the combined objective function.

(Snoek et al, 2012) using the package \texttt{scikit-optimize} (andreh7
et al, 2017) .

    \begin{Verbatim}[commandchars=\\\{\}]

        ---------------------------------------------------------------------------

        AttributeError                            Traceback (most recent call last)

        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
        271         try:
    --> 272             obj = getattr(obj, subpath)
        273         except AttributeError:

        AttributeError: 'module' object has no attribute 'objective'

    During handling of the above exception, another exception occurred:

        AttributeError                            Traceback (most recent call last)

        <ipython-input-8-3150368f60c1> in <module>()
          1 from skopt import load
          2 
    ----> 3 res\_gp = load(join(FIT\_DIR, 'additive\_cvge'))

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/skopt/utils.py in load(filename, **kwargs)
        122         Reconstructed OptimizeResult instance.
        123     """
    --> 124     return load\_(filename, **kwargs)

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in load(filename, mmap\_mode)
        573                     return load\_compatibility(fobj)
        574 
    --> 575                 obj = \_unpickle(fobj, filename, mmap\_mode)
        576 
        577     return obj

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in \_unpickle(fobj, filename, mmap\_mode)
        505     obj = None
        506     try:
    --> 507         obj = unpickler.load()
        508         if unpickler.compat\_mode:
        509             warnings.warn("The file '\%s' has been generated with a "

        /usr/lib/python3.4/pickle.py in load(self)
       1036                     raise EOFError
       1037                 assert isinstance(key, bytes\_types)
    -> 1038                 dispatch[key[0]](self)
       1039         except \_Stop as stopinst:
       1040             return stopinst.value

        /usr/lib/python3.4/pickle.py in load\_global(self)
       1323         module = self.readline()[:-1].decode("utf-8")
       1324         name = self.readline()[:-1].decode("utf-8")
    -> 1325         klass = self.find\_class(module, name)
       1326         self.append(klass)
       1327     dispatch[GLOBAL[0]] = load\_global

        /usr/lib/python3.4/pickle.py in find\_class(self, module, name)
       1375         \_\_import\_\_(module, level=0)
       1376         return \_getattribute(sys.modules[module], name,
    -> 1377                              allow\_qualname=self.proto >= 4)
       1378 
       1379     def load\_reduce(self):

        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
        273         except AttributeError:
        274             raise AttributeError("Can't get attribute \{!r\} on \{!r\}"
    --> 275                                  .format(name, obj))
        276     return obj
        277

        AttributeError: Can't get attribute 'objective' on <module '\_\_main\_\_'>

    \end{Verbatim}

    % Add a bibliography block to the postdoc

\bibliographystyle{unsrt}
\bibliography{refs}

    \end{document}

### 2017-04-24 _paper_output/paper_sparse_hawkes.tex
mysterious declaration randomisation
cites

@@
 
+\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
+\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
+\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
+\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
+\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
 \expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
-\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
+\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
-\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
-\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
-\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
-\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
-\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
 \expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
 \expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
-\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
-\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
 \expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
-\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
-\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
 
@@
 
-The Hawkes process (Hawkes, 1971) specifically, is the linear,
+The Hawkes process \cite{hawkes_point_1971} specifically, is the linear,
 self-exciting point process; that is, given a "background" rate
@@
 
-    One such assumption would be penalization :citeTikhonov1965,Hoerl1970
--{}-I apply an additional penalty to the log likelihood function,
-penalizing particular values of the parameters.
-
-For the penalized log likelihood for parameter $\theta$ and penalty
-$\pi$, I write
+    One such assumption would be penalization -{}-I apply an additional term
+to the log likelihood function, penalizing particular values of the
+parameters. For the penalized log likelihood for parameter $\theta$ and
+penalty $\pi$, I write
 
@@
 
-Hereafter, I will drop the prime from the augmented parameter vector and
-simply write $\theta$.
-
-Penalization is more frequently presented in the context of generalized
-linear regression from i.id. samples, but it also fits within a
-generalized Maximum Likelihood estimation theory.
-
-Various gradient ascent algorithms are known to perform well for
-:cite::Simon2011,Wu2008. Pragmatically, for performance reasons, I used
-a mixed strategy. My algorithm attempts to update marginal estimates for
-each $\omega_i$ parameter more rapidly first via Newton's method
-:citeBattiti1992,Ozaki1979 and uses conjugate gradient descent for the
-parametric estimates of the Hawkes parameters. If the updates steps are
-small this seems to be stable. There are various tuning parameters to
-such algorithms.
-
-Many variants of pathwise regularization and exist, such as backwards
-stepwise regularization, versions with random initialization, Least
-Angle Regression, :citeEfron2004a selection with integrated cross
-validation and so on. :citeFriedman2010
-
-Many variants of pathwise regularization and exist, such as backwards
-stepwise regularization, versions with random initialization, Least
-Angle Regression, :citeEfron2004a selection with integrated cross
-validation and so on. :citeFriedman2010 For this particular model I have
-found few results giving me analytic guarantees, so I use heuristics and
-simulation-based verification. I do not declare that this method "best"
-in any sense, or that it will find the correct global maximum penalized
-likelihood etc. The method, however, is simple enough to prototype
-rapidly and, as it turns out, performs well on test data. Results will
-be presented in the next chapter.
+For this particular model I have found few results giving me analytic
+guarantees, so I use heuristics and simulation-based verification. I do
+not declare that this method "best" in any sense, or that it will find
+the correct global maximum penalized likelihood etc. The method,
+however, is simple enough to prototype rapidly and, as it turns out,
+performs well on test data. Results will be presented in the next
+chapter.
 
@@
 computational and statistical convenience, including Laguerre
-polynomials (Ogata and Akaike, 1982), Exponential functions (Schoenberg
-2015; Rambaldi et al, 2015), and step functions (Eichler, 2016). As a
-convenient choice, we propose a basis of smooth unimodal non-negative
-functions with a computationally convenient form and a single mode, the
-convenient Maxwell distribution,
+polynomials \cite{ogata_linear_1982}, Exponential functions
+\cite{schoenberg_conistent_2005,rambaldi_modeling_2015}, and step
+functions \cite{eichler_graphical_2016} As a convenient choice, we
+propose a basis of smooth unimodal non-negative functions with a
+computationally convenient form and a single mode, the convenient
+Maxwell distribution,
 \[\phi_j(t):=\sqrt{\frac{2}{\pi}} \frac{t^2\exp\left(-\frac{t^2}{2\sigma_j^2}\right)}{\sigma_j^3}.\]
@@
 There is no closed-form solution for the optimum, but the derivatives of
-this function are easy to calculate manually see e.g. (Ozaki 1979). Here
-howevfer, we use automatic differentiation to make more complex
-penalisation procedure conveneient to differentiate (Baydin and
-Pearlmutter 2014).
+this function are easy to calculate manually see e.g.
+\cite{ozaki_maximum_1979}. Here howevfer, we use automatic
+differentiation to make more complex penalisation procedure conveneient
+to differentiate \cite{baydin_automatic_2014}.
 
@@
     To remedy this, I use Sugiura's finite-sample-corrected version, the
-\emph{AICc} (Sugiura, 1978).
+\emph{AICc} \cite{sugiura_further_1978}.
 \[{\mathrm {AIC}(L(X;\hat{\theta}))}=2d-2\ln(L)\]
@@
 The traditional means of performing model selection in Hawkes-type
-models e.g. Ogata (1983), uses \emph{degrees-of-freedom} penalties
-(Efron 1986). Specifically, the \emph{AIC}, the Akaike Information
-Criterion (Akaike 1981) is widely popular.
+models e.g. \cite{ogata_estimation_1983}, uses \emph{degrees-of-freedom}
+penalties \cite{efron_how_1886}. Specifically, the \emph{AIC}, the
+Akaike Information Criterion \cite{akaike_likelihood_1981} is widely
+popular.
 
@@
 dimensions, and in any case does not generalise readily to non-smooth
-penalties. (Konishi and kitagawa 2008)
+penalties. \cite{konishi_information_2008}
 
@@
 $\ell_1$ regression model provides an unbiassed estimate of the
-effective degrees of freedom of the model (Zou et al, 2007.)
+effective degrees of freedom of the model \cite{zou_degrees_2007}
 
@@
 
-(Snoek et al, 2012) using the package \texttt{scikit-optimize} (andreh7
-et al, 2017) .
+\cite{snoek_practical_2012} using the package \texttt{scikit-optimize}
+\cite{andreh7_scikit-optimize_2017} .
 
@@
 
-        <ipython-input-8-3150368f60c1> in <module>()
+        <ipython-input-12-3150368f60c1> in <module>()
           1 from skopt import load

### 2017-04-24 _paper_output/paper_sparse_hawkes.tex

@@
 
+\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
 \expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
-\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
+\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
 \expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
-\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
-\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
 \expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
-\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
 \expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
 \expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
-\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
-\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
-\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
-\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
-\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
-\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
-\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
 \expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
 
@@
 resembles a linear system, there are few of the conveneient results that
-woudl apply in a linear regression problem; and as a time series problem
+would apply in a linear regression problem; and as a time series problem
 it is difficult to choose our penalty parameters by standard
 non-parametric means such as cross validation. The proposed method uses
-a heuristically-motivated Akaike-style Information criterion
-regularisation parameter selection, which we test through simulation.
-The proposed method works well in practice.
+a heuristically-motivated degrees-of-freedom regularisation parameter
+selection, which we test through simulation. The proposed method works
+well in practice.
 
@@
 with respect to the filtration generated by $N$, i.e.
+
 \[\lambda(\mathcal{F(t^-)}):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
@@
 The Hawkes process \cite{hawkes_point_1971} specifically, is the linear,
-self-exciting point process; that is, given a "background" rate
+self-exciting point process; Concretely, given a "background" rate
 $\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,$ and a non-negative influence
@@
 
-In this problem, we are given a signal realisation of the count process
+In this problem, we are given a single realisation of the count process
 $N$ on an interval $T=[t_0,t_N]$. We are interested in identifying, from
 this data, two unknown functional parameters; both the background rate
-function $\mu$ and the influence kernel $\phi$.
+function $\mu$ and the influence kernel $\phi$. Since this is a general
+nonparametric problem we will use traditional regularisation techniques
+to control the expolosion of parameters.
 
-    One such assumption would be penalization -{}-I apply an additional term
-to the log likelihood function, penalizing particular values of the
-parameters. For the penalized log likelihood for parameter $\theta$ and
-penalty $\pi$, I write
+    We hope to recover the excitation kernel, which is unknown, and also a
+background signal, by appropriate penalisation of the terms.
+
+For the penalized log likelihood for parameter $\theta$ and penalty
+$\pi$, I write
 
@@
 penalizes only the non-parametric extensions to the model, such that
-they vanish as the penalty term increases. :citeGreen1987
+they vanish as the penalty term increases. \cite{green_penalized_1987}
 
@@
 $\omega(t)$ from the parametric estimate.
-
-For this particular model I have found few results giving me analytic
-guarantees, so I use heuristics and simulation-based verification. I do
-not declare that this method "best" in any sense, or that it will find
-the correct global maximum penalized likelihood etc. The method,
-however, is simple enough to prototype rapidly and, as it turns out,
-performs well on test data. Results will be presented in the next
-chapter.
 
@@
 polynomials \cite{ogata_linear_1982}, Exponential functions
-\cite{schoenberg_conistent_2005,rambaldi_modeling_2015}, and step
+\cite{schoenberg_consistent_2005,rambaldi_modeling_2015}, and step
 functions \cite{eichler_graphical_2016} As a convenient choice, we
@@
 
-    \subsubsection{Optimization}\label{optimization}
-
     The non-penalised log-likelihood function for a given realisation of the
-data is given
+data is given \cite{ozaki_etimating_1979}
 
@@
 
-And using the method of maximum likelihood we estimate our parameter
-vector by numerical optimisation of
+Using the method of maximum likelihood we estimate our parameter vector
+by numerical optimisation of
 \[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta).\]
 
-There is no closed-form solution for the optimum, but the derivatives of
-this function are easy to calculate manually see e.g.
-\cite{ozaki_maximum_1979}. Here howevfer, we use automatic
-differentiation to make more complex penalisation procedure conveneient
-to differentiate \cite{baydin_automatic_2014}.
-
-We are able to use these to devise rapid gradient-pased optimisation
-procedure.
+    There is no closed-form solution for the optimum of the Hawkes model
+likelihood, penalised or not. The derivatives of the unpenalised
+likelihood are easy to calculate manually -{}-{}- see e.g.
+\cite{ozaki_maximum_1979}. Here, however, we use automatic
+differentiation to make more complex penalisation procedure convenient
+to differentiate \cite{baydin_automatic_2014}. The optimisation proceeds
+using a standard black-box based Truncated Newton algorithm; as more
+specialised sparse regression algorithms such as Least-Angle Regression,
+do not directly apply here.
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_13_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_14_0.pdf}
     \end{center}

### 2017-04-24 branching_process/cts/background.py
yay count weighting

@@ -154,6 +154,10 @@ class StepKernel(ConstKernel):
         kappa = np.maximum(kappa, -mu)
         return np.ones_like(t) * (mu + np.amax(kappa))
 
+    def count(self, t, *args, **kwargs):
+        tau = self.get_param('tau', **kwargs)
+        return np.histogram(t, tau, density=False)
+
 
 class MultiplicativeStepKernel(StepKernel):
     """

### 2017-04-24 branching_process/cts/fit.py

@@ -113,16 +113,23 @@ class ContinuousExact(object):
     def penalty(
             self,
             kappa=None,
-            tau=None,
             omega=None,
             pi_kappa=0.0,
             pi_omega=0.0,
+            weight_count=None,
             **kwargs
             ):
         if kappa is None:
             kappa = self.phi_kernel.get_param('kappa')
         if omega is None:
             omega = self.mu_kernel.get_param('kappa')
+        if weight_count is None:
+            weight_count = self._weight_count
+
+        if weight_count:
+            weight = np.ones_like(omega)
+        else:
+            weight = self.mu_kernel.count(**kwargs)
 
         # 2 different l_1 penalties
         pen = 0
@@ -132,7 +139,7 @@ class ContinuousExact(object):
             )
         if self._fit_omega:
             pen = pen + np.sum(
-                np.abs(omega) * pi_omega
+                np.abs(omega) * pi_omega * weight
             )
         return pen
 
@@ -189,6 +196,7 @@ class ContinuousExact(object):
             t_end=None,
             param_vector=None,
             tol=None,
+            weight_count=False,
             **kwargs):
 
         self._t_start = t_start
@@ -244,6 +252,7 @@ class ContinuousExact(object):
         self._n_kappa_pack = self.n_phi_bases * self._fit_kappa
         self._n_omega_pack = self.n_mu_bases * self._fit_omega
         self._n_tau_pack = self.n_phi_bases * self._fit_tau
+        self._weight_count = weight_count
         self._mu_bounds = mu_kernel.mu_bounds()
         if self._fit_kappa:
             self._kappa_bounds = phi_kernel.kappa_bounds()

### 2017-04-24 branching_process/cts/influence.py

@@ -15,7 +15,8 @@ class InfluenceKernel(object):
             self,
             n_bases=1,
             eps=1e-8,
-            *args, **fixed_kwargs):
+            *args,
+            **fixed_kwargs):
         self._fixed_kwargs = fixed_kwargs
         self._fixed_kwargs.setdefault('kappa', np.ones(n_bases)/n_bases)
         self._fixed_kwargs.setdefault('tau', np.arange(n_bases)+1)

### 2017-04-25 branching_process/cts/fit.py
hack to avoid zero weight

@@ -126,10 +126,9 @@ class ContinuousExact(object):
         if weight_count is None:
             weight_count = self._weight_count
 
+        weight = np.ones_like(omega)
         if weight_count:
-            weight = np.ones_like(omega)
-        else:
-            weight = self.mu_kernel.count(**kwargs)
+            weight += self.mu_kernel.count(**kwargs)
 
         # 2 different l_1 penalties
         pen = 0

### 2017-04-25 branching_process/cts/fit.py
actually use weight without crashing

@@ -116,19 +116,12 @@ class ContinuousExact(object):
             omega=None,
             pi_kappa=0.0,
             pi_omega=0.0,
-            weight_count=None,
             **kwargs
             ):
         if kappa is None:
             kappa = self.phi_kernel.get_param('kappa')
         if omega is None:
             omega = self.mu_kernel.get_param('kappa')
-        if weight_count is None:
-            weight_count = self._weight_count
-
-        weight = np.ones_like(omega)
-        if weight_count:
-            weight += self.mu_kernel.count(**kwargs)
 
         # 2 different l_1 penalties
         pen = 0
@@ -138,7 +131,7 @@ class ContinuousExact(object):
             )
         if self._fit_omega:
             pen = pen + np.sum(
-                np.abs(omega) * pi_omega * weight
+                np.abs(omega) * pi_omega * self._omega_weight
             )
         return pen
 
@@ -271,6 +264,17 @@ class ContinuousExact(object):
             self._omega_bounds +
             self._tau_bounds
         )
+        if weight_count is None:
+            weight_count = self._weight_count
+
+        self._omega_weight = np.ones(self.n_mu_bases)
+
+        ## Nb this will break if we ever fit the background rate tau
+        if weight_count:
+            self._omega_weight += np.array(self.mu_kernel.count(
+                self._eval_ts,
+                **kwargs
+            ), dtype='float')
 
     def _teardown_graphs(self):
 
@@ -291,6 +295,7 @@ class ContinuousExact(object):
             t_end=None,
             tol=None,
             coordwise=False,
+            weight_count=False,
             **kwargs
             ):
         self._setup_graphs(

### 2017-04-25 _paper_output/paper_sparse_hawkes.tex
much texing

@@
                                 % normalem makes italics be italics, not underlines
+
+\usepackage{apacite}
 
@@
 
-\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
+\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
+\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
 \expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
+\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
+\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
+\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
+\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
 \expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 \expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
-\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
-\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
-\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
 \expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
-\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
-\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
-\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
 \expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
-\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
-\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
-\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
-\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
-\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
-\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
 
@@
 
-    \section{$\ell_1$-penalised non-stationary Hawkes branching
-process}\label{ellux5f1-penalised-non-stationary-hawkes-branching-process}
+    \section{\texorpdfstring{\(\ell_1\)-penalised non-stationary Hawkes
+branching
+process}{\textbackslash{}ell\_1-penalised non-stationary Hawkes branching process}}\label{ell_1-penalised-non-stationary-hawkes-branching-process}
 
@@
 parameters of a Hawkes point process and identifying a sparse driving
-signal by means of $\ell_1$ penalised regression where, by assumption,
+signal by means of \(\ell_1\) penalised regression where, by assumption,
 the driving noise has a different form. This is analaogous to a blind
-deconvolution problem or a compressive sensing-style inverse problem,
-but considerably more challenging; While the process model superfcially
-resembles a linear system, there are few of the conveneient results that
-would apply in a linear regression problem; and as a time series problem
-it is difficult to choose our penalty parameters by standard
-non-parametric means such as cross validation. The proposed method uses
-a heuristically-motivated degrees-of-freedom regularisation parameter
-selection, which we test through simulation. The proposed method works
-well in practice.
+deconvolution problem, robust regression, or a compressive sensing-style
+inverse problem, but considerably more challenging; While the process
+model superfcially resembles a linear system, there are few of the
+convenient results that would apply in a linear regression problem; and
+as a time series problem it is difficult to choose our penalty
+parameters by standard non-parametric means such as cross validation.
+The proposed method uses a heuristically-motivated degrees-of-freedom
+regularisation parameter selection, which we test through simulation.
+The proposed method works well in practice.
 
@@
 
-We define call a counting process $N$ indexed by $t$ a (TODO)
-predictable Poisson point process if it has an associated
-\emph{intensity} process $\lambda$, such that $\lambda$ is predictable
-with respect to the filtration generated by $N$, i.e.
+We are given a counting process \(N\) on \(\mathbb R\), and an interval
+\([0,T]\). (TODO clarify fit interval versus data interval)
+
+We write the time series either by its values
+\(\mathbf{t}=\{t_1, t_2,\dots,t_N(T)\}\). We represent the time series
+either by its values or its arrival times as notationally conveneient.
+We call it a (TODO) predictable Poisson point process if it has an
+associated \emph{intensity} process \(\lambda\), such that \(\lambda\)
+is predictable with respect to the filtration generated by \(N\), i.e.
+\cite{daley_introduction_2003}
 
@@
 
-The Hawkes process \cite{hawkes_point_1971} specifically, is the linear,
-self-exciting point process; Concretely, given a "background" rate
-$\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,$ and a non-negative influence
-kernel with positive support
-$\phi:\mathbb{R}^+\rightarrow\mathbb{R}^+\cap\{0\}$, the Hawkes process
+    The Hawkes process \cite{hawkes_point_1971} in particular, is the
+linear, self-exciting point process; Concretely, given a "background"
+rate \(\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,\) and a non-negative
+influence kernel with positive support
+\(\phi:\mathbb{R}\rightarrow\mathbb{R}^+\cap\{0\}\), the Hawkes process
 has the following expression for its intensity:
@@
 \begin{aligned}
-\lambda(\mathcal{F(t)}^-_{(-\infty,t)}) &= \mu(t) + \phi_\theta * N\\
-&= \mu(t)  + \int_{-\infty}^{\infty}\phi_\theta(t-s)dN(s)\\
-&= \mu(t) + \int_{-\infty}^{t}\phi_\theta(t-s)dN(s)\\
-&= \mu(t) + \sum_{t_i\lt t}\phi_\theta(t-t_i)
+\lambda(\mathcal{F(N(t))}^-_{(-\infty,t)}) &= \mu(t) + \phi * N\\
+&= \mu(t)  + \int_{-\infty}^{\infty}\phi(t-s)dN(s)\\
+&= \mu(t) + \int_{-\infty}^{t}\phi(t-s)dN(s)\\
+&= \mu(t) + \sum_{t_i\lt t}\phi(t-t_i)
 \end{aligned}
@@
 
-In this problem, we are given a single realisation of the count process
-$N$ on an interval $T=[t_0,t_N]$. We are interested in identifying, from
-this data, two unknown functional parameters; both the background rate
-function $\mu$ and the influence kernel $\phi$. Since this is a general
-nonparametric problem we will use traditional regularisation techniques
-to control the expolosion of parameters.
+For brevity, we write
+\(\lambda(\mathcal{F(t)}^-_{(-\infty,N(t))})=\lambda(t)\).
 
-    We hope to recover the excitation kernel, which is unknown, and also a
-background signal, by appropriate penalisation of the terms.
+    The log-likelihood function for a given realisation of the data is given
+\cite{ozaki_etimating_1979}
 
-For the penalized log likelihood for parameter $\theta$ and penalty
-$\pi$, I write
+\[L_\pi(N;\mu,\phi):=-\int_0^T\lambda(t;\mu,\phi)dt + \int_0^T\log \lambda(t;\mu,\phi) dN_t \]
 
-\[L_\pi(\mathbf t, \theta'):= L_\pi(\mathbf t, \theta')- \pi P(\theta')\]
+Using the method of maximum likelihood we estimate our parameter vector
+by numerical optimisation of
+\[\hat{\theta}_\pi(N) = \operatorname{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
 
-$P$ here is a non-negative functional which penalizes certain values of
-the parameter vector, and $\pi\geq 0$ is the penalty weight
-hyperparameter.
-
-As before, the estimate is defined as the maximizer:
-
-\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta)\]
-
-For non-parametric extension to a parametric model, one usually
-penalizes only the non-parametric extensions to the model, such that
-they vanish as the penalty term increases. \cite{green_penalized_1987}
-
-\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\mathbf t) =\hat{\theta}(\mathbf t)\]
-
-I follow this practice here, penalizing only the deviations of
-$\omega(t)$ from the parametric estimate.
-
-    We take $\mu$ to be an unknown sparse, non-random signal in a particular
-sense. We will assume that, given a linear basis
-$\{\psi_j \}_{1\leq j\leq M}$, that we may write the background rate as
-\[\mu(t) = \mu_0 + \sum_{j=1}^M  \omega_j \psi_j(t).\] \emph{Sparsity}
-here will imply that the coefficients $\omega_j$ will be mostly 0.
+    We take \(\mu\) to be an unknown sparse, non-random signal in a
+particular sense. We will assume that, given a basis
+\(\{\psi_j \}_{1\leq j\leq M}\), that we may write the background rate
+as \[\mu(t) = \mu_0 + \sum_{j=1}^M  \omega_j \psi_j(t).\]
+\emph{Sparsity} here will imply that the coefficients \(\omega_j\) will
+be mostly 0.
 
@@
 \[\psi_j(t) := \mathbb{I}_{[\tau_{j},\tau_{j+1})}(t)\] for some lattice
-$\{\tau_j\},$ which will correspond to the assumption that "sparse"
+\(\{\tau_j\},\) which will correspond to the assumption that "sparse"
 triggers will have localised support in time; we are looking for short
@@
 
-Similarly, we will approximate $\phi$ in a different functional basis
-${f}_{1\leq k \leq K}$, such that $\phi(s):= \sum_{k=1}^K f(s;k)$. This
-choice is likely arbitrary, but many exist in the literature trading off
-computational and statistical convenience, including Laguerre
-polynomials \cite{ogata_linear_1982}, Exponential functions
+Similarly, we will approximate \(\phi\) in a different functional basis
+\(\{\bar{\psi}\}_{1\leq k \leq K}\), such that
+\(\phi(t):= \sum_{k=1}^K \bar{\psi}(t;k)\). This basis choice is likely
+arbitrary, but many exist in the literature trading off computational
+and statistical convenience, including Laguerre polynomials
+\cite{ogata_linear_1982}, Exponential functions
 \cite{schoenberg_consistent_2005,rambaldi_modeling_2015}, and step
@@
 propose a basis of smooth unimodal non-negative functions with a
-computationally convenient form and a single mode, the convenient
-Maxwell distribution,
-\[\phi_j(t):=\sqrt{\frac{2}{\pi}} \frac{t^2\exp\left(-\frac{t^2}{2\sigma_j^2}\right)}{\sigma_j^3}.\]
+computationally convenient form, the Maxwell density,
+\[\bar{\psi}_j(t;\sigma):=\sqrt{\frac{2}{\pi}} \frac{t^2\exp\left(-\frac{t^2}{2\sigma^2}\right)}{\sigma^3}.\]
+so that
 
-A motivation for these specific choices is a particular real-world
-problem: modelling viral memes in social media. We suspect that online
-popularity is governed by some mixture of organic social sharing,
-modeled by the self-exciting $\phi$ component, mixed with intermittent
-external boosts, such as when an internet meme is featured on an
-external medium such as television.
-
-    mean delay $\mu=2\tau \sqrt{\frac{2}{\pi}}$
-
-mode $\sqrt{2}\tau$
-
-scale $\sigma=\sqrt{\frac{\tau^2(3 \pi - 8)}{\pi}}$
-
-We take fixed $M$-point lattice $\tilde{\tau}$ and $K$-fixed scales
-$\tilde{\sigma}.$
-
-The parameter vector to be estimated then is
-$\theta:=(\mu, \tilde{\omega}, \tilde{\kappa}).$
+\[\phi(t):= \sum_{j=1}^K \kappa_j\bar{\psi}(t;\sigma_j).\] Once again,
+some vector \(\{\sigma_j\}_{j\leq K}\) of time-scale parameters are
+chosen \emph{a priori}.
 
@@
 
-    The non-penalised log-likelihood function for a given realisation of the
-data is given \cite{ozaki_etimating_1979}
+    Since this is a general nonparametric problem we will use traditional
+regularisation techniques to control estimates. We hope to recover the
+unknown excitation kernel and a background intensity by appropriate
+penalisation of the terms.
 
-\[L_\pi(\mathbf t;\theta):=-\int_0^T\lambda_\theta(t)dt + \int_0^T\log \lambda_\theta(t) dN_t - \pi\|\boldsymbol \omega_\theta(t)\|_1\]
+For the penalized log likelihood for penalty functional \(\pi\), we
+write
 
-Using the method of maximum likelihood we estimate our parameter vector
-by numerical optimisation of
-\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta).\]
+\[L_\pi(N, \mu, \phi):= L(N; \mu, \phi)- \pi(\mu, \phi)\]
+
+As in plain log-likelihood optimization, the estimate is defined as the
+maximizer:
+
+\[\hat{\theta}_\pi(N) = \operatorname{argmax}_\theta L_\pi(N;\theta)\]
+
+    A motivation for these specific choices is a particular real-world
+problem: modelling viral sharing and spreading in social media. We
+suspect that online popularity is governed by some mixture of two
+componenets:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  organic social sharing, modeled by the self-exciting \(\phi\)
+  component, and
+\item
+  intermittent external boosts represented as variations in the the
+  \(\mu\) component, such as when an internet meme is featured on an
+  external medium such as television.
+\end{enumerate}
+
+Models such as the Hawkes process have frequently been applied in an
+attempt to disambiguate these two
+\cite{crane_robust_2008,mitchell_hawkes_2010,rizoiu_expecting_2017}.
+
+    \(\ell_1\) type penalties are known for their ability to produce sparse
+solutions, so we choose penalties of the form
+\[\pi(\theta)=p\left\|\{\sigma_j\}\right\|_1+q\left\|\{\omega_j\}\right\|_1.\]
+With the right choice of the parameters \(p\) and \(q\) we hope to
+shrink unnecessary parameters to zero, identifying predictively
+necessary ones, as in Lasso selection.
 
@@
 likelihood, penalised or not. The derivatives of the unpenalised
-likelihood are easy to calculate manually -{}-{}- see e.g.
+likelihood are easy to calculate manually -\/-\/-
 \cite{ozaki_maximum_1979}. Here, however, we use automatic
@@
 do not directly apply here.
-
-    To remedy this, I use Sugiura's finite-sample-corrected version, the
-\emph{AICc} \cite{sugiura_further_1978}.
-\[{\mathrm {AIC}(L(X;\hat{\theta}))}=2d-2\ln(L)\]
-
-\[{\mathrm {AICc}}:={\mathrm {AIC}}+{\frac {2d(d+1)}{N-d-1}}\]
-
-    \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_14_0.pdf}
-    \end{center}
-    { \hspace*{\fill} \\}
 
@@
 models e.g. \cite{ogata_estimation_1983}, uses \emph{degrees-of-freedom}
-penalties \cite{efron_how_1886}. Specifically, the \emph{AIC}, the
-Akaike Information Criterion \cite{akaike_likelihood_1981} is widely
-popular.
+penalties \cite{efron_how_1886}. Specifically, Akaike's
+\cite{akaike_likelihood_1981} Information Criterion, or \emph{AIC}, is
+widely popular.
 
-The AIC formula, for a model $M$ fit to a given data set $X$, for
-estimated parameter vector $\hat{\theta^M}$ with log likelihood $L$ and
-degrees of freedom $d^M$
+The AIC formula, for a given log-likelihood parameter estimate
+\(\hat{\theta}(N)\), with 'degrees of freedom', \(d\) is given
 
-\[\mathrm{AIC}(X,M)=2d^M-2L^M(X,\hat{\theta}^M)\]
+\[\mathrm{AIC}(N,\hat{\theta}(N))=2d-2L(N,\hat{\theta}(N))\]
 
-The degrees of freedom for many models taken to be equal to the
-cardinality of the parameter vector $\theta$.
+The degrees of freedom for parametric models is taken to be the number
+of free parameters, equal to the cardinality of the parameter vector
+\(\theta\).
 
-Given two ML fitted models, $(M_1,\hat{\theta}^M_1)$ and
-$(M_2\,\hat{\theta}^M_2)$, the difference in their AIC values is an
-estimate of the relative Kullback-Leibler divergence of the inferred
-measures $\hat{\mu_1},\hat{\mu_2}$ from the unknown true distribution,
-$\mu$ i.e.
+    The AIC does not immediately solve the problem of regularized parameter
+selection where the regularisation parameter may be a real value
+smoothly interpolating many possible models. One can calculate an AIC
+for penalised models, but this require onerous Hessian matrix inversions
+and becomes costly in high dimensions, and in any case does not
+generalise readily to non-smooth penalties.
+\cite{konishi_information_2008}
 
-\[\mathrm{AIC}(X,M_1)-\mathrm{AIC}(X,M_2) \simeq D_\mathrm{KL}(\mu\|\hat{\mu_1}) - D_\mathrm{KL}(\mu\|\hat{\mu_2})\]
+We proceed heuristically, based on the degrees-of-freedom arguments from
+Lasso-type linear-model regularisation. There, the number of non-zero
+coefficients fit under a particular value of the regularization penalty
+\(\pi\) in an \(\ell_1\) regression model provides an unbiassed estimate
+of the effective degrees of freedom of the model
+\cite{zou_degrees_2007}.
 
-However, the AIC does not immediately solve the problem of
-regularizition parameter selection in maximum likelihood, where the
-regularisation parameter may be a real value smoothly interpolating many
-possible models. One can calculate an AIC for penalised models, but this
-require onerous Hessian matrix inversions and becomes costly in high
-dimensions, and in any case does not generalise readily to non-smooth
-penalties. \cite{konishi_information_2008}
-
-We proceed heuristically, based on the arguments from Lasso-type
-linear-model regularisation. There, the number of non-zero coefficients
-fit under a particular value of the regularization penalty $\pi$ in an
-$\ell_1$ regression model provides an unbiassed estimate of the
-effective degrees of freedom of the model \cite{zou_degrees_2007}
-
-    This suggests an adjusted likelihood parameter of the following form:
+    This suggests, by analogy, an AIC estimate based on the number of
+non-zero parameters for given regularization coefficients:
 
 \[
-\widehat{\mathrm{AICc}}(X, \hat{\theta}_\pi) = {\frac {2Nd(\hat{\theta}_\pi) }{N-d(\hat{\theta}_\pi)-1}} - 2\ln(L(X;\hat{\theta}_\pi), X))
+\hat{d}\left(N,\hat{\theta},\pi\right)=\sum_j\mathbb 1\left\{\hat{\theta_j}(N,p,q)\neq 0\right\}
 \]
 
-where $d_\pi$ counts the number of non-zero coefficients estimated for
-given regularization coefficient $\pi$.
+    The suggested combined procedure is to combine, in an inner loop, the
+fitting of model parameters for given values of \(p\) and \(q\), and
+selecting the hyperparameters with the best overall AIC value.
+
+    \subsection{Simulations}\label{simulations}
+
+    We simulate data with parameters to match the stylized behaviour of the
+social media time series of previous work
+\cite{crane_robust_2008,mitchell_hawkes_2010,rizoiu_expecting_2017}, to
+test the efficacy of the proposed method.
+
+    \begin{center}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_0.pdf}
+    \end{center}
+    { \hspace*{\fill} \\}
+
+    \begin{center}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_1.pdf}
+    \end{center}
+    { \hspace*{\fill} \\}
+
+    \begin{center}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_2.pdf}
+    \end{center}
+    { \hspace*{\fill} \\}
 
     Extensive grid search for optimsal parameters is expensive, so we select
-both regularisation parameters using optimal sampling by Gaussian
-process regression of the combined objective function.
+both regularisation parameters using optimal sequential sampling by
+Gaussian process regression of the combined objective function
+\cite{snoek_practical_2012} using the software package
+\href{https://github.com/scikit-optimize/scikit-optimize}{scikit-optimize}.
 
-\cite{snoek_practical_2012} using the package \texttt{scikit-optimize}
-\cite{andreh7_scikit-optimize_2017} .
+    \subsection{Conclusions}\label{conclusions}
+
+    \subsection{Appendix and nonsense}\label{appendix-and-nonsense}
 
@@
 
-        AttributeError                            Traceback (most recent call last)
+        FileNotFoundError                         Traceback (most recent call last)
 
-        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
-        271         try:
-    --> 272             obj = getattr(obj, subpath)
-        273         except AttributeError:
-
-        AttributeError: 'module' object has no attribute 'objective'
-
-    During handling of the above exception, another exception occurred:
-
-        AttributeError                            Traceback (most recent call last)
-
-        <ipython-input-12-3150368f60c1> in <module>()
+        <ipython-input-8-765ef5deb338> in <module>()
           1 from skopt import load
           2 
-    ----> 3 res\_gp = load(join(FIT\_DIR, 'additive\_cvge'))
+    ----> 3 res\_gp = load(join(FIT\_DIR, 'additive\_cvge.gp.bz2'))
 
-        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/skopt/utils.py in load(filename, **kwargs)
+        /usr/local/lib/python3.6/site-packages/skopt/utils.py in load(filename, **kwargs)
         122         Reconstructed OptimizeResult instance.
@@
 
-        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in load(filename, mmap\_mode)
-        573                     return load\_compatibility(fobj)
-        574 
-    --> 575                 obj = \_unpickle(fobj, filename, mmap\_mode)
-        576 
-        577     return obj
+        /usr/local/lib/python3.6/site-packages/sklearn/externals/joblib/numpy\_pickle.py in load(filename, mmap\_mode)
+        565             obj = \_unpickle(fobj)
+        566     else:
+    --> 567         with open(filename, 'rb') as f:
+        568             with \_read\_fileobject(f, filename, mmap\_mode) as fobj:
+        569                 if isinstance(fobj, \_basestring):
 
-        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in \_unpickle(fobj, filename, mmap\_mode)
-        505     obj = None
-        506     try:
-    --> 507         obj = unpickler.load()
-        508         if unpickler.compat\_mode:
-        509             warnings.warn("The file '\%s' has been generated with a "
-
-        /usr/lib/python3.4/pickle.py in load(self)
-       1036                     raise EOFError
-       1037                 assert isinstance(key, bytes\_types)
-    -> 1038                 dispatch[key[0]](self)
-       1039         except \_Stop as stopinst:
-       1040             return stopinst.value
-
-        /usr/lib/python3.4/pickle.py in load\_global(self)
-       1323         module = self.readline()[:-1].decode("utf-8")
-       1324         name = self.readline()[:-1].decode("utf-8")
-    -> 1325         klass = self.find\_class(module, name)
-       1326         self.append(klass)
-       1327     dispatch[GLOBAL[0]] = load\_global
-
-        /usr/lib/python3.4/pickle.py in find\_class(self, module, name)
-       1375         \_\_import\_\_(module, level=0)
-       1376         return \_getattribute(sys.modules[module], name,
-    -> 1377                              allow\_qualname=self.proto >= 4)
-       1378 
-       1379     def load\_reduce(self):
-
-        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
-        273         except AttributeError:
-        274             raise AttributeError("Can't get attribute \{!r\} on \{!r\}"
-    --> 275                                  .format(name, obj))
-        276     return obj
-        277
-
-        AttributeError: Can't get attribute 'objective' on <module '\_\_main\_\_'>
+        FileNotFoundError: [Errno 2] No such file or directory: '/Users/dan/Datasets/branching\_process\_output/additive\_cvge.gp.bz2'
 
@@
 
+    Given two ML fitted models, \((M_1,\hat{\theta}^M_1)\) and
+\((M_2\,\hat{\theta}^M_2)\), the difference in their AIC values is an
+estimate of the relative Kullback-Leibler divergence of the inferred
+measures \(\hat{\mu_1},\hat{\mu_2}\) from the unknown true distribution,
+\(\mu\) i.e.
+
+\[\mathrm{AIC}(X,M_1)-\mathrm{AIC}(X,M_2) \simeq D_\mathrm{KL}(\mu\|\hat{\mu_1}) - D_\mathrm{KL}(\mu\|\hat{\mu_2})\]
+
+    mean delay \(\mu=2\tau \sqrt{\frac{2}{\pi}}\)
+
+mode \(\sqrt{2}\tau\)
+
+scale \(\sigma=\sqrt{\frac{\tau^2(3 \pi - 8)}{\pi}}\)
+
+We take fixed \(M\)-point lattice \(\tilde{\tau}\) and \(K\)-fixed
+scales \(\tilde{\sigma}.\)
+
+The parameter vector to be estimated then is
+\(\theta:=(\mu, \tilde{\omega}, \tilde{\kappa}).\)
+
+    The AIC is based on asymptotic arguments, and Sugiura's
+finite-sample-corrected version \emph{AICc}, has lower bias for most
+practical applications \cite{sugiura_further_1978}.
+
+\[{\mathrm {AICc}}:={\mathrm {AIC}}+{\frac {2d(d+1)}{N_T-d-1}}\]
+
+    For non-parametric extension to a parametric model, one usually
+penalizes only the non-parametric extensions to the model, such that
+they vanish as the penalty term increases. \cite{green_penalized_1987}
+
+\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\mathbf t) =\hat{\theta}(\mathbf t)\]
+
+I follow this practice here, penalizing only the deviations of
+\(\omega(t)\) from the parametric estimate.
+
+\[L_\pi(\mathbf t;\mu,\phi):=-\int_0^T\lambda(t;\mu,\phi)dt + \int_0^T\log \lambda(t;\mu,\phi) dN_t - \pi\|\boldsymbol \omega_\theta(t)\|_1\]
+
     % Add a bibliography block to the postdoc
 
-\bibliographystyle{unsrt}
+\bibliographystyle{apacite}
 \bibliography{refs}

### 2017-04-26 _paper_output/paper_sparse_hawkes.tex
make everything nice and explained

@@
 likelihood are easy to calculate manually -\/-\/-
-\cite{ozaki_maximum_1979}. Here, however, we use automatic
-differentiation to make more complex penalisation procedure convenient
-to differentiate \cite{baydin_automatic_2014}. The optimisation proceeds
-using a standard black-box based Truncated Newton algorithm; as more
-specialised sparse regression algorithms such as Least-Angle Regression,
-do not directly apply here.
+\cite{ozaki_maximum_1979}. Here, however, we use standard reverse-mode
+automatic differentiation \cite{baydin_automatic_2014}. The optimisation
+proceeds using a standard gradient-based black-box based Truncated
+Newton algorithm; as more specialised sparse regression algorithms such
+as Least-Angle Regression, do not directly apply here.
 
@@
 
-The degrees of freedom for parametric models is taken to be the number
+The degrees of freedom for a parametric model is taken to be the number
 of free parameters, equal to the cardinality of the parameter vector
@@
 coefficients fit under a particular value of the regularization penalty
-\(\pi\) in an \(\ell_1\) regression model provides an unbiassed estimate
+\(\pi\) in an \(\ell_1\) regression model provides an unbiased estimate
 of the effective degrees of freedom of the model
@@
 \[
-\hat{d}\left(N,\hat{\theta},\pi\right)=\sum_j\mathbb 1\left\{\hat{\theta_j}(N,p,q)\neq 0\right\}
+\hat{d}\left(N,\hat{\theta},p,q\right)=\sum_j\mathbb 1\left\{\hat{\theta_j}(N,p,q)\neq 0\right\}
 \]
@@
 
+    TODO: switch to consistent step-function display of instantaneous rate.
+
     We simulate data with parameters to match the stylized behaviour of the
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_1.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_1.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_22_2.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_2.pdf}
     \end{center}
@@
 
+\texttt{\color{outcolor}Out[{\color{outcolor}27}]:}
+
+    \begin{center}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_25_0.png}
+    \end{center}
+    { \hspace*{\fill} \\}
+
+    With a single regularisation parameter, only for the background rate
+coefficient \(\omega_j\), we recover the background rate efficiently.
+
+\texttt{\color{outcolor}Out[{\color{outcolor}28}]:}
+
+    \begin{center}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_27_0.png}
+    \end{center}
+    { \hspace*{\fill} \\}
+
+    \subsection{How to simulate a Hawkes
+process}\label{how-to-simulate-a-hawkes-process}
+
+Here we follow a standard "thinning" method of simulating Hawkes
+processes, given as Algorithm 7.5.IV of Daley and Vere-Jones'
+\emph{Introduction to the Theory of Point Processes Volume 1}
+\cite{daley_introduction_2003}, restating Ogata's original Algorithm 2
+(\cite{ogata_lewis_1981}.
+
+Since that is a general case, here we present a specific version for
+linear Hawkes processes which is considerably simpler.
+
+For clarity, we will use the representation of the Hawkes process as an
+immigrant-birth process, where the recursive "self-excited" component
+corresponds to the "births". We will assume that we are given a list of
+arrival times of immigrants, \(\{a_1,a_2,\dots,a_N\}\), and wish to
+simulate the births arising from these immigrants. We will associate a
+point and its arrival time and discuss them interchangeably where
+ambiguity does not arise.
+
+Each of these \(a_i\) gives rise to a family of descendants
+\(\mathrm{descendants}(a_i)\), distributed independently of all other
+families, whose combined superposition is the final Hawkes process
+(\cite{hawkes_cluster_1974}. It will suffice to simulate the arrival
+times of each of these families. Without loss of generality we may
+consider the problem of simulating the births in a single family started
+at time \(a_1=0\) and translate it.
+
+Now the families
+\(\mathrm{descendants}(t_j)\forall t_j\in \mathrm{children}(a_i)\) are
+all once again identically and mutually independently distributed.
+Proceeding by induction, it suffices to simulate the children of a
+single point at time \(0\). The same method will suffice for its
+children, children of children and so on. The final realisation can be
+constructed from these simualtions by appropriate concatenation and
+translation of the generations of each family.
+
+The single kernel simulation is simple, since the Poisson intensity
+process is no longer conditional on the history of the sequence. As
+before, the influence kernel is \(\phi\). We will assume that \(\phi\)
+is a.e. continuous \(L_1\) integrable and eventually decreasing in
+\(t\). We associate with \(\phi\) a "majorant", \(\phi^{**}\) such that
+\[\phi^{**}(t)\geq\phi(t), \forall t>0\] and \(\phi^{**}\) is
+non-increasing and integrable. With the unimodal influence kernels used
+in the current paper, such majorants are trivial to construct.
+
+We will use the fact that there is some time \(T>0\) such that we do not
+consider events after it, i.e. that we only care about arrival times on
+\([0,T]\), and so only children that occur before time \(T\) can be
+relevent.
+
+\subsubsection{Algorithm: Ogata's thinning
+algorithm}\label{algorithm-ogatas-thinning-algorithm}
+
+Our goal is to simulate the set \(S=\mathrm{children}(a_0)\).
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Set \(i=1, t=0\), \(S=\varnothing\)
+\item
+  Generate \(\Delta\sim \operatorname{Exp}(1/\phi^{**}(t))\)
+\item
+  Set \(t\leftarrow t+\Delta\)
+\item
+  If \(t\geq T\), halt.
+\item
+  With probability \(P=\frac{\phi(t)}{\phi^{**}(t)}\), set
+  \(S\leftarrow S\cup \{t\}\).
+\item
+  Continue from step 2.
+\end{enumerate}
+
     \subsection{Conclusions}\label{conclusions}
@@
     \subsection{Appendix and nonsense}\label{appendix-and-nonsense}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-        ---------------------------------------------------------------------------
-
-        FileNotFoundError                         Traceback (most recent call last)
-
-        <ipython-input-8-765ef5deb338> in <module>()
-          1 from skopt import load
-          2 
-    ----> 3 res\_gp = load(join(FIT\_DIR, 'additive\_cvge.gp.bz2'))
-
-        /usr/local/lib/python3.6/site-packages/skopt/utils.py in load(filename, **kwargs)
-        122         Reconstructed OptimizeResult instance.
-        123     """
-    --> 124     return load\_(filename, **kwargs)
-
-        /usr/local/lib/python3.6/site-packages/sklearn/externals/joblib/numpy\_pickle.py in load(filename, mmap\_mode)
-        565             obj = \_unpickle(fobj)
-        566     else:
-    --> 567         with open(filename, 'rb') as f:
-        568             with \_read\_fileobject(f, filename, mmap\_mode) as fobj:
-        569                 if isinstance(fobj, \_basestring):
-
-        FileNotFoundError: [Errno 2] No such file or directory: '/Users/dan/Datasets/branching\_process\_output/additive\_cvge.gp.bz2'
-
-    \end{Verbatim}
 

### 2017-04-26 dodo.py

@@ -173,6 +173,10 @@ def task_view_pdf_paper_sparse_hawkes():
     return _task_view_pdf('paper_sparse_hawkes')
 
 
+def task_zdravko_paper_sparse_hawkes():
+    return _task_zdravko('paper_sparse_hawkes', 'sparse_hawkes')
+
+
 def task_html_paper_sparse_hawkes():
     return _task_html('paper_sparse_hawkes')
 

### 2017-04-27 _paper_output/paper_sparse_hawkes.tex
Explicate

@@
 We are interested in the problem of simultaneous identification of the
-parameters of a Hawkes point process and identifying a sparse driving
-signal by means of \(\ell_1\) penalised regression where, by assumption,
-the driving noise has a different form. This is analaogous to a blind
-deconvolution problem, robust regression, or a compressive sensing-style
-inverse problem, but considerably more challenging; While the process
-model superfcially resembles a linear system, there are few of the
-convenient results that would apply in a linear regression problem; and
-as a time series problem it is difficult to choose our penalty
-parameters by standard non-parametric means such as cross validation.
-The proposed method uses a heuristically-motivated degrees-of-freedom
-regularisation parameter selection, which we test through simulation.
-The proposed method works well in practice.
+parameters of a Hawkes point process and identifying an inhomogeneous
+driving signal by means of \(\ell_1\) penalised regression. This is
+analaogous to a blind deconvolution problem \cite{stockham_blind_1975},
+adaptive robust regression \cite{donoho_automatic_1988}, or a sparse
+inverse problem \cite{tibshirani_regression_1996}, or, but considerably
+more challenging; While the process model superficially resembles a
+linear system, there are few of the convenient results that would apply
+in a linear regression problem; and as a time series problem it is
+difficult to choose our penalty parameters by standard non-parametric
+means such as cross validation. The proposed method uses a
+heuristically-motivated degrees-of-freedom regularisation parameter
+selection, which we test through simulation. The proposed method works
+well in practice.
 
     \subsection{Introduction}\label{introduction}
+
+\subsubsection{Notation}\label{notation}
 
@@
 
-We write the time series either by its values
-\(\mathbf{t}=\{t_1, t_2,\dots,t_N(T)\}\). We represent the time series
-either by its values or its arrival times as notationally conveneient.
+We represent a realisation of the time series by its arrival times
+\(\mathbf{t}=\{t_1, t_2,\dots,t_N(T)\}\). We may also, when as
+notationally conveneient, use the equivalent representation as the
+function that counts the arrivals:
+
+\[
+N(t):= \left|\{s:s\in \mathbf{t}\,\&\,s\leq t\}\right|
+\]
+
+We use only the left-continuous filtration
+\(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the process, which we
+here write as \(\mathcal{F}(N)\)
+
 We call it a (TODO) predictable Poisson point process if it has an
 associated \emph{intensity} process \(\lambda\), such that \(\lambda\)
-is predictable with respect to the filtration generated by \(N\), i.e.
+is predictable with respect to \(\mathcal{F}(N)\) generated by the
+process, i.e.
+
+\[\lambda(\mathcal{F}(t)):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
 \cite{daley_introduction_2003}
-
-\[\lambda(\mathcal{F(t^-)}):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
 
@@
 influence kernel with positive support
-\(\phi:\mathbb{R}\rightarrow\mathbb{R}^+\cap\{0\}\), the Hawkes process
+\(\phi:\mathbb{R}\rightarrow\mathbb{R}^+\cup\{0\}\), the Hawkes process
 has the following expression for its intensity:
@@
 Here we follow a standard "thinning" method of simulating Hawkes
-processes, given as Algorithm 7.5.IV of Daley and Vere-Jones'
+processes, specialised from Algorithm 7.5.IV of Daley and Vere-Jones'
 \emph{Introduction to the Theory of Point Processes Volume 1}
@@
 
-Since that is a general case, here we present a specific version for
-linear Hawkes processes which is considerably simpler.
+Since that is a general case, here we present, informally, a simpler,
+specific version for linear Hawkes processes.
 
@@
 children, children of children and so on. The final realisation can be
-constructed from these simualtions by appropriate concatenation and
+constructed from these simulations by appropriate concatenation and
 translation of the generations of each family.
@@
 \[\phi^{**}(t)\geq\phi(t), \forall t>0\] and \(\phi^{**}\) is
-non-increasing and integrable. With the unimodal influence kernels used
-in the current paper, such majorants are trivial to construct.
+non-increasing and integrable. As an aside, we mention that for the
+unimodal influence kernels used in the current paper, such majorants are
+trivial to construct.
 

### 2017-04-27 dodo.py
it's a chapter not a paper

@@ -162,23 +162,23 @@ def _task_zdravko(srcpth, destpth):
 
 
 def task_latex_paper_sparse_hawkes():
-    return _task_latex('paper_sparse_hawkes')
+    return _task_latex('chapter_sparse_hawkes')
 
 
 def task_pdf_paper_sparse_hawkes():
-    return _task_pdf('paper_sparse_hawkes')
+    return _task_pdf('chapter_sparse_hawkes')
 
 
 def task_view_pdf_paper_sparse_hawkes():
-    return _task_view_pdf('paper_sparse_hawkes')
+    return _task_view_pdf('chapter_sparse_hawkes')
 
 
 def task_zdravko_paper_sparse_hawkes():
-    return _task_zdravko('paper_sparse_hawkes', 'sparse_hawkes')
+    return _task_zdravko('chapter_sparse_hawkes', 'sparse_hawkes')
 
 
 def task_html_paper_sparse_hawkes():
-    return _task_html('paper_sparse_hawkes')
+    return _task_html('chapter_sparse_hawkes')
 
 
 def task_html_intro_to_cts_hawkes():

### 2017-04-27 dodo.py
rerename

@@ -161,23 +161,23 @@ def _task_zdravko(srcpth, destpth):
     )
 
 
-def task_latex_paper_sparse_hawkes():
+def task_latex_chapter_sparse_hawkes():
     return _task_latex('chapter_sparse_hawkes')
 
 
-def task_pdf_paper_sparse_hawkes():
+def task_pdf_chapter_sparse_hawkes():
     return _task_pdf('chapter_sparse_hawkes')
 
 
-def task_view_pdf_paper_sparse_hawkes():
+def task_view_pdf_chapter_sparse_hawkes():
     return _task_view_pdf('chapter_sparse_hawkes')
 
 
-def task_zdravko_paper_sparse_hawkes():
+def task_zdravko_chapter_sparse_hawkes():
     return _task_zdravko('chapter_sparse_hawkes', 'sparse_hawkes')
 
 
-def task_html_paper_sparse_hawkes():
+def task_html_chapter_sparse_hawkes():
     return _task_html('chapter_sparse_hawkes')
 
 

### 2017-04-28 _paper_output/chapter_sparse_hawkes.tex
notation updates

--- 2017-04-27 _paper_output/paper_sparse_hawkes.tex
@@
 We are interested in the problem of simultaneous identification of the
-parameters of a Hawkes point process and identifying an inhomogeneous
-driving signal by means of \(\ell_1\) penalised regression. This is
-analaogous to a blind deconvolution problem \cite{stockham_blind_1975},
-adaptive robust regression \cite{donoho_automatic_1988}, or a sparse
-inverse problem \cite{tibshirani_regression_1996}, or, but considerably
-more challenging; While the process model superficially resembles a
-linear system, there are few of the convenient results that would apply
-in a linear regression problem; and as a time series problem it is
-difficult to choose our penalty parameters by standard non-parametric
-means such as cross validation. The proposed method uses a
-heuristically-motivated degrees-of-freedom regularisation parameter
-selection, which we test through simulation. The proposed method works
-well in practice.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  self excitation kernel, and
+\item
+  (possibly inhomogenous) background rate
+\end{enumerate}
+
+of a Hawkes point process a using of \(\ell_1\) penalised regression.
+
+This is analaogous to a blind deconvolution problem
+\cite{stockham_blind_1975}, adaptive robust regression
+\cite{donoho_automatic_1988}, or a sparse inverse problem
+\cite{tibshirani_regression_1996}, or, but considerably more
+challenging; While the process model superficially resembles a linear
+system, there are few of the convenient results that would apply in a
+linear regression problem; and as a time series problem it is difficult
+to choose our penalty parameters by standard non-parametric means such
+as cross validation. The proposed method uses a heuristically-motivated
+degrees-of-freedom regularisation parameter selection, which we test
+through simulation. The proposed method works well in practice.
 
@@
 
-We are given a counting process \(N\) on \(\mathbb R\), and an interval
-\([0,T]\). (TODO clarify fit interval versus data interval)
+We are given a counting process \(N(t)\) on \(\mathbb R\)
 
-We represent a realisation of the time series by its arrival times
-\(\mathbf{t}=\{t_1, t_2,\dots,t_N(T)\}\). We may also, when as
-notationally conveneient, use the equivalent representation as the
-function that counts the arrivals:
+We represent a realisation of the time series by a (finite by
+assumption) sequence of arrival times
+\(\mathbf{t}=\{\dots,t_1, t_2,\dots,t_N(T)\}.\) We will be concerned
+with estimating its likelihood based on the likelihood of the series on
+an interval an interval \([0,T]\). We define
+\(\mathbf{t}':=\mathbf{t}\cap [0,T]\) and assume
+\(\left|\mathbf{t}'\right|<\infty.\)
 
-\[
-N(t):= \left|\{s:s\in \mathbf{t}\,\&\,s\leq t\}\right|
+We also, when notationally convenient, use the equivalent representation
+as the left-continuous function that increments at each arrival time: \[
+N(t)-N(s):= \left|\{s:s\in \mathbf{t}\,\mathrm{and}\,s\leq t\}\right| \,\mathrm{ for }\, t\geq s
 \]
+
+We set \(\lim_{s\rightarrow 0} N(s)\equiv 0\) and
+\(t_1\equiv \inf \mathbf{t}'\); Indices \(i\leq 0\) are reserved for
+point in the "history" of the process before time \(0\), which may be
+generated by any arbitrary process. We will not be concerned with
+evaluating the likelihood of such histories but may retain them to
+control for their influence on the future of the series. For the rates
+to be well-defined, however, we will assume that the increment sof the
+process are finite on any finite interval.
 
@@
 \(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the process, which we
-here write as \(\mathcal{F}(N)\)
+here write as \(\mathcal{F}(N(t)).\)
 
-We call it a (TODO) predictable Poisson point process if it has an
-associated \emph{intensity} process \(\lambda\), such that \(\lambda\)
-is predictable with respect to \(\mathcal{F}(N)\) generated by the
-process, i.e.
+We call it a regular point process if it has an associated
+\emph{intensity} process \(\lambda\), such that \(\lambda\) is
+predictable with respect to \(\mathcal{F}(N)\) generated by the process,
+i.e. \cite{daley_introduction_2003}
 
-\[\lambda(\mathcal{F}(t)):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
-\cite{daley_introduction_2003}
+\[\lambda(t|\mathcal{F}(N(t))):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
 
-    The Hawkes process \cite{hawkes_point_1971} in particular, is the
-linear, self-exciting point process; Concretely, given a "background"
-rate \(\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,\) and a non-negative
+Hereafter we define \(\lambda^*(t):=\lambda(t|\mathcal{F}(N(t)))\) for
+brevity.
+
+Such regular point processes generalise the homogeneous Poisson rate in
+the sense that if we choose \(\lambda*(t)\equiv \lambda_0\) then it is
+the intensity of a homogeneous Poisson rate with fixed intensity
+\(\lambda_0\).
+
+    \[N(t)\sim \operatorname{Pois}(\Lambda(t|\mathcal{F}(N(t)))\] where
+\(\Lambda(s,t|\mathcal{F}(N(t)))=\int_s^t \lambda(s|\mathcal{F}(N(s))).\)
+
+The waiting times are exponentially distributed given the intensity,
+i.e. their densities \(f_i(s)\) are given \[\begin{aligned}
+f_i(t)&\equiv f(t_{i+1}-t_i|\mathcal{F}(N(t_i)))\\
+&=\Lambda(t_i,t|\mathcal{F}(N(t_i))\exp\left(-t\Lambda(t_i,t|\mathcal{F}(N(t_i))\right)
+\end{aligned}\]
+
+We will use the for the inter-arrival waiting times \(s_i=t_{i+1}-t_i\)
+to calculate the data likelihood, after Rubin \cite{rubin_regular_1972}.
+
+    The waiting times are exponentially distributed given the intensity,
+i.e. their densities \(f_i(s)\) for any given transition are given
+\[\begin{aligned}
+\frac{d}{dt}P(t_{i+1}=t|\mathcal{F}(N(t)))
+\end{aligned}\] fa \[\begin{aligned}
+f_i(t)&\equiv f(t_{i+1}-t_i|\mathcal{F}(N(t_i)))\\
+&=\Lambda(t_i,t|\mathcal{F}(N(t_i))\exp\left(-t\Lambda(t_i,t|\mathcal{F}(N(t_i))\right)
+\end{aligned}\]
+
+    It follows that the log likelihood \(\ell_i:=\log f_i\) of such a
+waiting time \(s_i\) is given
+
+\[\begin{aligned}
+\ell_i(t_{i+1}-t_i) &=\log \Lambda(s_|\mathcal{F}(N(t_i)) - s_i \Lambda(s_i|\mathcal{F}(N(t_i))\\
+&= \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds
+\end{aligned}
+\]
+
+    We may therefore write the likelihood for the data \[
+\begin{aligned}
+\ell(\mathbf{t}) &= \sum_{1\leq i\leq N(T)}\ell_i(t_{i+1}-t_i)\\
+&= \sum_{1\leq i\leq N(T)} \left( \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds\right)\\
+&= \sum_{1\leq i\leq N(T)} \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - \sum_{1\leq i\leq N(T)} {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds\\
+&= \sum_{1\leq i\leq N(T)} \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - \sum_{1\leq i\leq N(T)} {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds
+\end{aligned}
+\]
+
+    The Hawkes process \cite{hawkes_point_1971} in particular, is the linear
+self-exciting point process; Concretely, given a "background" rate
+\(\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,\) and a non-negative
 influence kernel with positive support
@@
 \begin{aligned}
-\lambda(\mathcal{F(N(t))}^-_{(-\infty,t)}) &= \mu(t) + \phi * N\\
+\lambda(t|\mathcal{F}(N(t);\mu,\phi) &= \mu(t) + \phi * N\\
 &= \mu(t)  + \int_{-\infty}^{\infty}\phi(t-s)dN(s)\\
@@
 For brevity, we write
-\(\lambda(\mathcal{F(t)}^-_{(-\infty,N(t))})=\lambda(t)\).
+\(\lambda(\mathcal{F}(N(t));\mu,\phi)=\lambda^*(t)\), but will suppress
+explicity dependence on history and parameters..
 
-    The log-likelihood function for a given realisation of the data is given
+    The log-likelihood function for a given realisation of the data is can
+be found from the usual likelihood for Poisson data of a given intensity
 \cite{ozaki_etimating_1979}
 
-\[L_\pi(N;\mu,\phi):=-\int_0^T\lambda(t;\mu,\phi)dt + \int_0^T\log \lambda(t;\mu,\phi) dN_t \]
+\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda^*(t)dt + \int_0^T\log \lambda^*(t) dN_t \\
+&=-\int_0^T\lambda^*(t)dt + \int_0^T\log \lambda^*(t) dN_t \\
+&=-\int_0^T\lambda^*(t)dt + \sum_{t\in\mathbf{t}\cap [0,T]}\log \lambda^*(t)
+\end{aligned}\]
 
 Using the method of maximum likelihood we estimate our parameter vector
-by numerical optimisation of
+\(\theta:=(\mu, \phi)\) by choosing the estimates to maximise the
+likelihood,
 \[\hat{\theta}_\pi(N) = \operatorname{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
+
+    Here we extend the likelihood approach, treating it as penalised
+functional regression problem. We augment the log-likeihood function
+with a penalty functional \(\pi\):
+
+\[L_\pi(N, \mu, \phi):= L(N; \mu, \phi)- \pi(\mu, \phi)\]
+
+As in plain log-likelihood optimization, the estimate is defined as the
+maximizer:
+
+\[\hat{\theta}_\pi(N) = \operatorname{argmax}_\theta L_\pi(N;\theta)\]
 
@@
 and statistical convenience, including Laguerre polynomials
-\cite{ogata_linear_1982}, Exponential functions
+\cite{ogata_linear_1982}, exponential functions
 \cite{schoenberg_consistent_2005,rambaldi_modeling_2015}, and step
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_10_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_15_0.pdf}
     \end{center}
     { \hspace*{\fill} \\}
-
-    Since this is a general nonparametric problem we will use traditional
-regularisation techniques to control estimates. We hope to recover the
-unknown excitation kernel and a background intensity by appropriate
-penalisation of the terms.
-
-For the penalized log likelihood for penalty functional \(\pi\), we
-write
-
-\[L_\pi(N, \mu, \phi):= L(N; \mu, \phi)- \pi(\mu, \phi)\]
-
-As in plain log-likelihood optimization, the estimate is defined as the
-maximizer:
-
-\[\hat{\theta}_\pi(N) = \operatorname{argmax}_\theta L_\pi(N;\theta)\]
 
@@
 
+    With basis functions fixed, the problem has become one of selection of
+intercepts and basis expansions coefficients, and we may equivalently
+write
+\(\theta=(\mu, \kappa_1,\kappa_2,\dots,\kappa_K,\omega_1, \omega_2,\dots,\omega_M).\)
+
 \(\ell_1\) type penalties are known for their ability to produce sparse
 solutions, so we choose penalties of the form
-\[\pi(\theta)=p\left\|\{\sigma_j\}\right\|_1+q\left\|\{\omega_j\}\right\|_1.\]
+\[\pi(\theta)=p\left\|\{\kappa_j\}\right\|_1+q\left\|\{\omega_j\}\right\|_1.\]
 With the right choice of the parameters \(p\) and \(q\) we hope to
@@
     There is no closed-form solution for the optimum of the Hawkes model
-likelihood, penalised or not. The derivatives of the unpenalised
-likelihood are easy to calculate manually -\/-\/-
-\cite{ozaki_maximum_1979}. Here, however, we use standard reverse-mode
-automatic differentiation \cite{baydin_automatic_2014}. The optimisation
-proceeds using a standard gradient-based black-box based Truncated
-Newton algorithm; as more specialised sparse regression algorithms such
-as Least-Angle Regression, do not directly apply here.
+likelihood, penalised or not. However, we may use the gradient to find
+it efficiently numerically. The optimisation proceeds using a standard
+gradient-based black-box based Truncated Newton algorithm; as more
+specialised sparse regression algorithms such as Least-Angle Regression,
+do not directly apply here.
+
+The gradient is efficient to calculate using standard standard
+reverse-mode automatic differentiation \cite{baydin_automatic_2014} for
+a wide variety of functional decomposition bases and penalties.
+
+The enthusiast may also calculate the derivatives analytically. For
+example, Ozaki \cite{ozaki_maximum_1979} give derivatives for the
+parametric log likelihood estimator.
+
+    Let \(\gamma\) be any of the components of the parameter vector
+\(\theta\). Since the likelihood depends on this vector only through the
+intensity process, we have
+
+\[\begin{aligned}
+{ \scriptstyle \frac{\partial}{\partial\gamma} } L_\pi(N;\theta)&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)dt + \int_0^T { \scriptstyle \frac{\partial}{\partial\gamma} }\log \lambda^*(t;\theta) dN_t -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta) \\
+&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)dt + \sum_{0\leq t_i\leq T} \frac{{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t_i)}{\lambda^*(t_i)} -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta)
+\end{aligned}
+\]
+
+    Taking \(\gamma=\mu\),
+
+\[{ \scriptstyle \frac{\partial}{\partial\mu} }\lambda^*(t;\mu) = 1\]
+
+I use higher derivatives for \(\mu\) so that I may optimize this
+component using a higher order Newton method, since we know that its
+optimum is typically shallow \cite{veen_estimation_2008}.
+
+\[\begin{aligned}
+  { \scriptstyle \frac{\partial}{\partial\mu} } L_\pi(N;\theta) &=-T + \sum_{0\leq t_i\leq T} \frac{1}{\lambda(t_i;\theta)} \\
+  { \scriptstyle \frac{\partial^2}{\partial\mu^2} } L_\pi(N;\theta) &= \sum_{0\leq t_i\leq T} \frac{-1}{\lambda^2(t_i;\theta)}\\
+  { \scriptstyle \frac{\partial^3}{\partial\mu^3} } L_\pi(N;\theta) &= \sum_{0\leq t_i\leq T} \frac{2}{\lambda^3(t_i;\theta)}
+  \end{aligned}\]
+
+    Now I handle \(\gamma=\kappa_j\),
+
+\[{ \scriptstyle \frac{\partial}{\partial\kappa_j} }\lambda^*(t;\theta) =  \sum_{t_i\lt t}\phi_j(t-t_i)\]
+
+so that
+
+\[\begin{aligned}
+  { \scriptstyle \frac{\partial}{\partial\kappa_i} } L_\pi(N; \theta)&=-\int_0^T \sum_{t_i\lt t}\phi_i(t-t_i) dt + \sum_{0\leq t_i\leq T} \frac{\sum_{t_k\lt t_i}\phi_j(t_i-t_k)}{\lambda_\theta(t_i)} -\frac{\partial}{\partial\kappa_j}\pi(\theta)
+\end{aligned}\]
+
+    Finally, I handle the \(\omega_j\) values.
+
+Taking \(\theta=\omega_j\), and defining
+\(\Delta\tau_j:= {\tau_{j-1}}-{\tau_j}\) we find
+
+\[
+\begin{align*}
+{ \scriptstyle \frac{\partial}{\partial\omega_j} }\pi\|\boldsymbol \omega_\theta(t)\|_1 &= { \scriptstyle \frac{\partial}{\partial\omega_j} }\pi \int_{\tau_{j-1}}^{\tau_j} \left|\omega_j\right|dt\\
+&=\pi\Delta\tau_j\operatorname{sgn} \omega_j\\
+{ \scriptstyle \frac{\partial}{\partial\omega_j} }\lambda_\theta(t|\mathcal{F}_t) &=  \Delta\tau_j\mathbb{I}_{[\tau_{j-1},\tau_j)}(t)\\
+{ \scriptstyle \frac{\partial}{\partial\omega_j} } L_\pi(\mathbf t;\theta) &=-\Delta\tau_j-\pi\Delta\tau_j\operatorname{sgn} \omega_j + \sum_{\tau_{j-1}\leq t_i\leq \tau_j} \frac{1}{\lambda_\theta(t_i)}
+\end{align*}
+\]
+
+Higher partial derivatives are also analogous to the partial derivatives
+with respect to \(\mu\), although of course the penalty introduces a
+discontinuity at \(\omega_j=0\). This last formula is the key to the
+gradient ascent algorithm.
+
+Note that the \(\omega_j\) values are mutually orthogonal, in the sense
+that \({ \scriptstyle \frac{\partial^2}{\partial\omega_i\omega_j}} =0\)
+if \(i\ne j\). I can treat these components, more or less, as separate
+univariate components when optimizing, and the Hessian will be sparse
+off the diagonal.
+
+By construction,
+\({ \scriptstyle \frac{\partial}{\partial\mu} }\pi(\gamma) = { \scriptstyle \frac{\partial}{\partial\eta} }\pi(\gamma) = { \scriptstyle \frac{\partial}{\partial\kappa} }\pi(\gamma) = 0\).
+
+TODO: I've changed \(\omega\) penalisation; this needs fixing.
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_1.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_1.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_23_2.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_2.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_25_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_33_0.png}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_27_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_35_0.png}
     \end{center}

### 2017-05-01 _paper_output/chapter_sparse_hawkes.tex
more Zifying

@@
 
-This is analaogous to a blind deconvolution problem
+This is analagous to a blind deconvolution problem
 \cite{stockham_blind_1975}, adaptive robust regression
 \cite{donoho_automatic_1988}, or a sparse inverse problem
-\cite{tibshirani_regression_1996}, or, but considerably more
-challenging; While the process model superficially resembles a linear
-system, there are few of the convenient results that would apply in a
-linear regression problem; and as a time series problem it is difficult
-to choose our penalty parameters by standard non-parametric means such
-as cross validation. The proposed method uses a heuristically-motivated
-degrees-of-freedom regularisation parameter selection, which we test
-through simulation. The proposed method works well in practice.
+\cite{tibshirani_regression_1996}, but considerably more challenging;
+While the process model superficially resembles an i.i.d. linear
+process, the different noise model and sequential dependence mean that
+there are few of the same convenient results; We cannot easily, for
+example, choose our our penalty parameters by standard cross validation.
+The proposed method uses a heuristically-motivated degrees-of-freedom
+regularisation parameter selection, which we test through simulation.
+The proposed method works well in practice.
 
@@
 
-We are given a counting process \(N(t)\) on \(\mathbb R\)
+We are given a process
+\(N:\mathcal{B}(\mathbb{R})\mapsto \mathbb{N}^0\), which counts the
+number of arrivals in a given time. Since we are usually concerned with
+arrivals over increasing time intervals, we will usually consider \(N\)
+to be indexed by time: \(N(t):=N([0,t))-N((-\infty, t))\).
 
@@
 
-We also, when notationally convenient, use the equivalent representation
-as the left-continuous function that increments at each arrival time: \[
-N(t)-N(s):= \left|\{s:s\in \mathbf{t}\,\mathrm{and}\,s\leq t\}\right| \,\mathrm{ for }\, t\geq s
-\]
-
-We set \(\lim_{s\rightarrow 0} N(s)\equiv 0\) and
-\(t_1\equiv \inf \mathbf{t}'\); Indices \(i\leq 0\) are reserved for
-point in the "history" of the process before time \(0\), which may be
-generated by any arbitrary process. We will not be concerned with
+We set \(t_1\equiv \inf \mathbf{t}'\); Indices \(i\leq 0\) are reserved
+for point in the "history" of the process before time \(0\), which may
+be generated by any arbitrary process. We will not be concerned with
 evaluating the likelihood of such histories but may retain them to
 control for their influence on the future of the series. For the rates
-to be well-defined, however, we will assume that the increment sof the
-process are finite on any finite interval.
+to be well-defined, however, we will assume that the increments of the
+process are a.s. finite on any finite interval.
 
@@
 
-\[\lambda(t|\mathcal{F}(N(t))):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]
+\[\lambda(t|\mathcal{F}(N(t))):=\lim_{h\to 0} \frac{\mathrm E\left(N([t,t+h))\right)}{h}.\]
 
@@
 Such regular point processes generalise the homogeneous Poisson rate in
-the sense that if we choose \(\lambda*(t)\equiv \lambda_0\) then it is
+the sense that if we choose \(\lambda^*(t)\equiv \lambda_0,\) then it is
 the intensity of a homogeneous Poisson rate with fixed intensity
@@
 
-    \[N(t)\sim \operatorname{Pois}(\Lambda(t|\mathcal{F}(N(t)))\] where
-\(\Lambda(s,t|\mathcal{F}(N(t)))=\int_s^t \lambda(s|\mathcal{F}(N(s))).\)
+    We will use the inter-arrival waiting times \(s_i=t_{i+1}-t_i\) to
+calculate the data likelihood, after Rubin \cite{rubin_regular_1972}.
 
-The waiting times are exponentially distributed given the intensity,
-i.e. their densities \(f_i(s)\) are given \[\begin{aligned}
-f_i(t)&\equiv f(t_{i+1}-t_i|\mathcal{F}(N(t_i)))\\
-&=\Lambda(t_i,t|\mathcal{F}(N(t_i))\exp\left(-t\Lambda(t_i,t|\mathcal{F}(N(t_i))\right)
+From the definition of the intensity we find \[\begin{aligned}
+\frac{d}{dt}P(t_{i+1}=t|\mathcal{F}(N(t)))=
 \end{aligned}\]
 
-We will use the for the inter-arrival waiting times \(s_i=t_{i+1}-t_i\)
-to calculate the data likelihood, after Rubin \cite{rubin_regular_1972}.
-
-    The waiting times are exponentially distributed given the intensity,
-i.e. their densities \(f_i(s)\) for any given transition are given
-\[\begin{aligned}
-\frac{d}{dt}P(t_{i+1}=t|\mathcal{F}(N(t)))
-\end{aligned}\] fa \[\begin{aligned}
-f_i(t)&\equiv f(t_{i+1}-t_i|\mathcal{F}(N(t_i)))\\
-&=\Lambda(t_i,t|\mathcal{F}(N(t_i))\exp\left(-t\Lambda(t_i,t|\mathcal{F}(N(t_i))\right)
-\end{aligned}\]
-
-    It follows that the log likelihood \(\ell_i:=\log f_i\) of such a
-waiting time \(s_i\) is given
-
-\[\begin{aligned}
-\ell_i(t_{i+1}-t_i) &=\log \Lambda(s_|\mathcal{F}(N(t_i)) - s_i \Lambda(s_i|\mathcal{F}(N(t_i))\\
-&= \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds
-\end{aligned}
-\]
-
-    We may therefore write the likelihood for the data \[
-\begin{aligned}
-\ell(\mathbf{t}) &= \sum_{1\leq i\leq N(T)}\ell_i(t_{i+1}-t_i)\\
-&= \sum_{1\leq i\leq N(T)} \left( \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds\right)\\
-&= \sum_{1\leq i\leq N(T)} \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - \sum_{1\leq i\leq N(T)} {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds\\
-&= \sum_{1\leq i\leq N(T)} \log \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds - \sum_{1\leq i\leq N(T)} {t_{i+1}} \int_{t_i}^{t_{i+1}} \lambda(s|\mathcal{F}(N(t_i))) ds
-\end{aligned}
-\]
+TODO: Which bits does Z want proven? Rewrite to start from transition
+probability? Probably makes this shorter. Otherwise prove equivalence of
+intensity and transition prob DE?
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_15_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_13_0.pdf}
     \end{center}
@@
 
-\[
-\begin{align*}
+\[\begin{aligned}
 { \scriptstyle \frac{\partial}{\partial\omega_j} }\pi\|\boldsymbol \omega_\theta(t)\|_1 &= { \scriptstyle \frac{\partial}{\partial\omega_j} }\pi \int_{\tau_{j-1}}^{\tau_j} \left|\omega_j\right|dt\\
@@
 { \scriptstyle \frac{\partial}{\partial\omega_j} } L_\pi(\mathbf t;\theta) &=-\Delta\tau_j-\pi\Delta\tau_j\operatorname{sgn} \omega_j + \sum_{\tau_{j-1}\leq t_i\leq \tau_j} \frac{1}{\lambda_\theta(t_i)}
-\end{align*}
-\]
+\end{aligned}\]
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_1.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_1.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_2.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_2.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_33_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_0.png}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_35_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_33_0.png}
     \end{center}
@@
 \cite{daley_introduction_2003}, restating Ogata's original Algorithm 2
-(\cite{ogata_lewis_1981}.
-
-Since that is a general case, here we present, informally, a simpler,
+(\cite{ogata_lewis_1981}. Here we present, informally, a simpler,
 specific version for linear Hawkes processes.
@@
 corresponds to the "births". We will assume that we are given a list of
-arrival times of immigrants, \(\{a_1,a_2,\dots,a_N\}\), and wish to
-simulate the births arising from these immigrants. We will associate a
-point and its arrival time and discuss them interchangeably where
-ambiguity does not arise.
+arrival times of immigrants, \(A=\{a_1,a_2,\dots,a_N\}\) from a
+realisation of the immigrant process. In this case, that immigrant
+process is, specifically, the inhomogenous Poisson process with
+unconditional intensity \(\mu(t)\). We associate a birth process \(X_i\)
+with each immigrant \(a_i\). Each such process is distributed
+independently of all other families, whose combined superposition is the
+final Hawkes process (\cite{hawkes_cluster_1974}.
 
-Each of these \(a_i\) gives rise to a family of descendants
-\(\mathrm{descendants}(a_i)\), distributed independently of all other
-families, whose combined superposition is the final Hawkes process
-(\cite{hawkes_cluster_1974}. It will suffice to simulate the arrival
-times of each of these families. Without loss of generality we may
-consider the problem of simulating the births in a single family started
-at time \(a_1=0\) and translate it.
+The clusters are generated as follows. For each \(a_i \in A\),
+independently generate children \(a_{i1},a_{i2}\dots\) according to a
+Poisson process with intensity \(\phi(t-a_i)\). Then, each child
+\(a_{ij}\) of \(a_i\) again generates a Poisson process with intensity
+\(\phi(t − a_{ij})\), and so on. The collection of all generated points
+from all descendants forms the Hawkes process.
 
-Now the families
-\(\mathrm{descendants}(t_j)\forall t_j\in \mathrm{children}(a_i)\) are
-all once again identically and mutually independently distributed.
-Proceeding by induction, it suffices to simulate the children of a
-single point at time \(0\). The same method will suffice for its
-children, children of children and so on. The final realisation can be
-constructed from these simulations by appropriate concatenation and
-translation of the generations of each family.
-
-The single kernel simulation is simple, since the Poisson intensity
-process is no longer conditional on the history of the sequence. As
-before, the influence kernel is \(\phi\). We will assume that \(\phi\)
-is a.e. continuous \(L_1\) integrable and eventually decreasing in
-\(t\). We associate with \(\phi\) a "majorant", \(\phi^{**}\) such that
+It suffices to know how to simulate a Poisson process with inhomogenous
+rate \(\phi\) started any point \(t_i\). We will assume that \(\phi\) is
+a.e. continuous \(L_1\) integrable and eventually decreasing in \(t\).
+We associate with \(\phi\) a "majorant", \(\phi^{**}\) such that
 \[\phi^{**}(t)\geq\phi(t), \forall t>0\] and \(\phi^{**}\) is
@@
 
-Our goal is to simulate the set \(S=\mathrm{children}(a_0)\).
+Our goal is to simulate the set \(S=\mathrm{children}(a_\chi)\).
 
@@
 \item
-  Set \(i=1, t=0\), \(S=\varnothing\)
+  Set \(i=1, t=t_i\), \(S=\varnothing\)
 \item
@@
     \subsection{Conclusions}\label{conclusions}
+
+    TODO
 
@@
 
+    The waiting times are exponentially distributed given the intensity,
+i.e. their densities \(f_i(s)\) are given \[\begin{aligned}
+f_i(t)&\equiv f(t_{i+1}-t_i|\mathcal{F}(N(t_i)))\\
+&=\Lambda(t_i,t|\mathcal{F}(N(t_i))\exp\left(-t\Lambda(t_i,t|\mathcal{F}(N(t_i))\right)
+\end{aligned}\]
+
+    We also, when notationally convenient, use the equivalent representation
+as the left-continuous function that increments at each arrival time: \[
+N(t)-N(s):= \left|\{s:s\in \mathbf{t}\,\mathrm{and}\,s\leq t\}\right| \,\mathrm{ for }\, t\geq s
+\]
+
+    \[N(t)\sim \operatorname{Pois}(\Lambda(t|\mathcal{F}(N(t)))\] where
+\(\Lambda(s,t|\mathcal{F}(N(t)))=\int_s^t \lambda(s|\mathcal{F}(N(s))).\)
+
     % Add a bibliography block to the postdoc

### 2017-05-01 _paper_output/chapter_sparse_hawkes.tex
macros

@@
 
+\newcommand{\var}{\operatorname{Var}}
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\bb}[1]{\mathbb{#1}}
+\renewcommand{\bf}[1]{\mathbf{#1}}
+\renewcommand{\v}[1]{\boldsymbol{#1}}
+\newcommand{\m}[1]{\mathrm{#1}}
+\newcommand{\cc}[1]{\mathcal{#1}}
+\renewcommand{\cc}[1]{\mathcal{#1}}
+\newcommand{\oo}[1]{\operatorname{#1}}
+
     % Prevent overflowing lines due to hard-to-break entities
@@
 process}{\textbackslash{}ell\_1-penalised non-stationary Hawkes branching process}}\label{ell_1-penalised-non-stationary-hawkes-branching-process}
+
+    \[
+\renewcommand{\var}{\operatorname{Var}}
+\renewcommand{\dd}{\mathrm{d}}
+\renewcommand{\bb}[1]{\mathbb{#1}}
+\renewcommand{\bf}[1]{\mathbf{#1}}
+\renewcommand{\v}[1]{\boldsymbol{#1}}
+\renewcommand{\m}[1]{\mathrm{#1}}
+\renewcommand{\cc}[1]{\mathcal{#1}}
+\renewcommand{\cc}[1]{\mathcal{#1}}
+\renewcommand{\oo}[1]{\operatorname{#1}}
+\]
 
@@
 
-We are given a process
-\(N:\mathcal{B}(\mathbb{R})\mapsto \mathbb{N}^0\), which counts the
-number of arrivals in a given time. Since we are usually concerned with
-arrivals over increasing time intervals, we will usually consider \(N\)
-to be indexed by time: \(N(t):=N([0,t))-N((-\infty, t))\).
+We are given a process \(N:\cc{B}(\bb{R})\mapsto \bb{N}^0\), which
+counts the number of arrivals in a given time. Since we are usually
+concerned with arrivals over increasing time intervals, we will usually
+consider \(N\) to be indexed by time:
+\(N(t):=N([0,t))-N((-\infty, t))\).
 
@@
 assumption) sequence of arrival times
-\(\mathbf{t}=\{\dots,t_1, t_2,\dots,t_N(T)\}.\) We will be concerned
-with estimating its likelihood based on the likelihood of the series on
-an interval an interval \([0,T]\). We define
-\(\mathbf{t}':=\mathbf{t}\cap [0,T]\) and assume
-\(\left|\mathbf{t}'\right|<\infty.\)
+\(\v{t}=\{\dots,t_1, t_2,\dots,t_N(T)\}.\) We will be concerned with
+estimating its likelihood based on the likelihood of the series on an
+interval an interval \([0,T]\). We define \(\v{t}':=\v{t}\cap [0,T]\)
+and assume \(\left|\v{t}'\right|<\infty.\)
 
-We set \(t_1\equiv \inf \mathbf{t}'\); Indices \(i\leq 0\) are reserved
-for point in the "history" of the process before time \(0\), which may
-be generated by any arbitrary process. We will not be concerned with
+We set \(t_1\equiv \inf \v{t}'\); Indices \(i\leq 0\) are reserved for
+point in the "history" of the process before time \(0\), which may be
+generated by any arbitrary process. We will not be concerned with
 evaluating the likelihood of such histories but may retain them to
@@
     We use only the left-continuous filtration
-\(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the process, which we
-here write as \(\mathcal{F}(N(t)).\)
+\(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the paths of the process,
+which we here write as \(\cc{F}(N(t)).\)
 
@@
 \emph{intensity} process \(\lambda\), such that \(\lambda\) is
-predictable with respect to \(\mathcal{F}(N)\) generated by the process,
-i.e. \cite{daley_introduction_2003}
+predictable with respect to \(\cc{F}(N)\) generated by the process, i.e.
+\cite{daley_introduction_2003}
 
-\[\lambda(t|\mathcal{F}(N(t))):=\lim_{h\to 0} \frac{\mathrm E\left(N([t,t+h))\right)}{h}.\]
+\[\lambda(t|\cc{F}(N(t))):=\lim_{h\to 0} \frac{\bb E\left(N([t,t+h))\right)}{h}.\]
 
-Hereafter we define \(\lambda^*(t):=\lambda(t|\mathcal{F}(N(t)))\) for
+Hereafter we define \(\lambda^*(t):=\lambda(t|\cc{F}(N(t)))\) for
 brevity.
@@
 From the definition of the intensity we find \[\begin{aligned}
-\frac{d}{dt}P(t_{i+1}=t|\mathcal{F}(N(t)))=
+\frac{\dd}{\dd t}\bb P(t_{i+1}=t|\mathcal{F}(N(t)))=
 \end{aligned}\]
@@
 \lambda(t|\mathcal{F}(N(t);\mu,\phi) &= \mu(t) + \phi * N\\
-&= \mu(t)  + \int_{-\infty}^{\infty}\phi(t-s)dN(s)\\
-&= \mu(t) + \int_{-\infty}^{t}\phi(t-s)dN(s)\\
+&= \mu(t)  + \int_{-\infty}^{\infty}\phi(t-s)\dd N(s)\\
+&= \mu(t) + \int_{-\infty}^{t}\phi(t-s)\dd N(s)\\
 &= \mu(t) + \sum_{t_i\lt t}\phi(t-t_i)
@@
 
-\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda^*(t)dt + \int_0^T\log \lambda^*(t) dN_t \\
-&=-\int_0^T\lambda^*(t)dt + \int_0^T\log \lambda^*(t) dN_t \\
-&=-\int_0^T\lambda^*(t)dt + \sum_{t\in\mathbf{t}\cap [0,T]}\log \lambda^*(t)
+\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda^*(t) \dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
+&=-\int_0^T\lambda^*(t)\dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
+&=-\int_0^T\lambda^*(t)\dd t + \sum_{t\in\v{t}'}\log \lambda^*(t)
 \end{aligned}\]
@@
 \(\theta:=(\mu, \phi)\) by choosing the estimates to maximise the
-likelihood,
-\[\hat{\theta}_\pi(N) = \operatorname{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
+log-likelihood,
+\[\hat{\theta}_\pi(N) = \o{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
 
-    Here we extend the likelihood approach, treating it as penalised
-functional regression problem. We augment the log-likeihood function
-with a penalty functional \(\pi\):
+    Here we extend the likelihood approach with a penalty functional
+\(\pi\):
 
@@
 
-\[\hat{\theta}_\pi(N) = \operatorname{argmax}_\theta L_\pi(N;\theta)\]
+\[\hat{\theta}_\pi(N) = \oo{argmax}_\theta L_\pi(N;\theta)\]
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_13_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_14_0.pdf}
     \end{center}
@@
 \[\begin{aligned}
-{ \scriptstyle \frac{\partial}{\partial\gamma} } L_\pi(N;\theta)&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)dt + \int_0^T { \scriptstyle \frac{\partial}{\partial\gamma} }\log \lambda^*(t;\theta) dN_t -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta) \\
-&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)dt + \sum_{0\leq t_i\leq T} \frac{{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t_i)}{\lambda^*(t_i)} -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta)
+{ \scriptstyle \frac{\partial}{\partial\gamma} } L_\pi(N;\theta)&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)dt + \int_0^T { \scriptstyle \frac{\partial}{\partial\gamma} }\log \lambda^*(t;\theta) \dd N_t -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta) \\
+&=-\int_0^T{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t;\theta)\dd t + \sum_{0\leq t_i\leq T} \frac{{ \scriptstyle \frac{\partial}{\partial\gamma} }\lambda^*(t_i)}{\lambda^*(t_i)} -{ \scriptstyle \frac{\partial}{\partial\gamma} }\pi(\theta)
 \end{aligned}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_1.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_1.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_29_2.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_2.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_31_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_32_0.png}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_33_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_34_0.png}
     \end{center}

### 2017-05-03 _paper_output/chapter_sparse_hawkes.tex
refs

@@
 be found from the usual likelihood for Poisson data of a given intensity
-\cite{ozaki_etimating_1979}
+\cite{ozaki_estimating_1979}
 
@@
 log-likelihood,
-\[\hat{\theta}_\pi(N) = \o{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
+\[\hat{\theta}_\pi(N) = \oo{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
 
@@
 In practice, we choose
-\[\psi_j(t) := \mathbb{I}_{[\tau_{j},\tau_{j+1})}(t)\] for some lattice
+\[\psi_j(t) := \bb{I}_{[\tau_{j},\tau_{j+1})}(t)\] for some lattice
 \(\{\tau_j\},\) which will correspond to the assumption that "sparse"
@@
 \[\begin{aligned}
-  { \scriptstyle \frac{\partial}{\partial\kappa_i} } L_\pi(N; \theta)&=-\int_0^T \sum_{t_i\lt t}\phi_i(t-t_i) dt + \sum_{0\leq t_i\leq T} \frac{\sum_{t_k\lt t_i}\phi_j(t_i-t_k)}{\lambda_\theta(t_i)} -\frac{\partial}{\partial\kappa_j}\pi(\theta)
+  { \scriptstyle \frac{\partial}{\partial\kappa_i} } L_\pi(N; \theta)&=-\int_0^T \sum_{t_i\lt t}\phi_i(t-t_i) \dd t + \sum_{0\leq t_i\leq T} \frac{\sum_{t_k\lt t_i}\phi_j(t_i-t_k)}{\lambda_\theta(t_i)} -\frac{\partial}{\partial\kappa_j}\pi(\theta)
 \end{aligned}\]
@@
 \[\begin{aligned}
-{ \scriptstyle \frac{\partial}{\partial\omega_j} }\pi\|\boldsymbol \omega_\theta(t)\|_1 &= { \scriptstyle \frac{\partial}{\partial\omega_j} }\pi \int_{\tau_{j-1}}^{\tau_j} \left|\omega_j\right|dt\\
-&=\pi\Delta\tau_j\operatorname{sgn} \omega_j\\
-{ \scriptstyle \frac{\partial}{\partial\omega_j} }\lambda_\theta(t|\mathcal{F}_t) &=  \Delta\tau_j\mathbb{I}_{[\tau_{j-1},\tau_j)}(t)\\
-{ \scriptstyle \frac{\partial}{\partial\omega_j} } L_\pi(\mathbf t;\theta) &=-\Delta\tau_j-\pi\Delta\tau_j\operatorname{sgn} \omega_j + \sum_{\tau_{j-1}\leq t_i\leq \tau_j} \frac{1}{\lambda_\theta(t_i)}
+{ \scriptstyle \frac{\partial}{\partial\omega_j} }\pi\|\v \omega_\theta(t)\|_1 &= { \scriptstyle \frac{\partial}{\partial\omega_j} }\pi \int_{\tau_{j-1}}^{\tau_j} \left|\omega_j\right|\dd t\\
+&=\pi\Delta\tau_j\oo{sgn} \omega_j\\
+{ \scriptstyle \frac{\partial}{\partial\omega_j} }\lambda_\theta(t|\mathcal{F}_t) &=  \Delta\tau_j\bb{I}_{[\tau_{j-1},\tau_j)}(t)\\
+{ \scriptstyle \frac{\partial}{\partial\omega_j} } L_\pi(\v t;\theta) &=-\Delta\tau_j-\pi\Delta\tau_j\oo{sgn} \omega_j + \sum_{\tau_{j-1}\leq t_i\leq \tau_j} \frac{1}{\lambda_\theta(t_i)}
 \end{aligned}\]
@@
 models e.g. \cite{ogata_estimation_1983}, uses \emph{degrees-of-freedom}
-penalties \cite{efron_how_1886}. Specifically, Akaike's
+penalties \cite{efron_how_1986}. Specifically, Akaike's
 \cite{akaike_likelihood_1981} Information Criterion, or \emph{AIC}, is
@@
 
-\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\mathbf t) =\hat{\theta}(\mathbf t)\]
+\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\v t) =\hat{\theta}(\v t)\]
 
@@
 
-\[L_\pi(\mathbf t;\mu,\phi):=-\int_0^T\lambda(t;\mu,\phi)dt + \int_0^T\log \lambda(t;\mu,\phi) dN_t - \pi\|\boldsymbol \omega_\theta(t)\|_1\]
+\[L_\pi(\v t;\mu,\phi):=-\int_0^T\lambda(t;\mu,\phi) \dd t + \int_0^T\log \lambda(t;\mu,\phi) \dd N_t - \pi\|\v \omega_\theta(t)\|_1\]
 
@@
 as the left-continuous function that increments at each arrival time: \[
-N(t)-N(s):= \left|\{s:s\in \mathbf{t}\,\mathrm{and}\,s\leq t\}\right| \,\mathrm{ for }\, t\geq s
+N(t)-N(s):= \left|\{s:s\in \v{t}\,\text{ and }\,s\leq t\}\right| \,\mathrm{ for }\, t\geq s
 \]

### 2017-05-08 _paper_output/chapter_sparse_hawkes.tex
fix up some typos

@@
 process, the different noise model and sequential dependence mean that
-there are few of the same convenient results; We cannot easily, for
-example, choose our our penalty parameters by standard cross validation.
-The proposed method uses a heuristically-motivated degrees-of-freedom
+there are fewer convenient results; We cannot easily, for example,
+choose our our penalty parameters by standard cross validation. The
+proposed method uses a heuristically-motivated degrees-of-freedom
 regularisation parameter selection, which we test through simulation.
@@
 counts the number of arrivals in a given time. Since we are usually
-concerned with arrivals over increasing time intervals, we will usually
-consider \(N\) to be indexed by time:
-\(N(t):=N([0,t))-N((-\infty, t))\).
+concerned with arrivals over simple intervals, we will understand that
+if \(N\) is indexed by a single point parameter, then it corresponds to
+the following counting function: \(N(t):=N([0,t))-N((-\infty, t))\).
 
@@
 \(\v{t}=\{\dots,t_1, t_2,\dots,t_N(T)\}.\) We will be concerned with
-estimating its likelihood based on the likelihood of the series on an
-interval an interval \([0,T]\). We define \(\v{t}':=\v{t}\cap [0,T]\)
-and assume \(\left|\v{t}'\right|<\infty.\)
+evaluating its likelihood at the arrivals in time \([0,T]\) We define
+\(\v{t}':=\v{t}\cap [0,T]\) and assume \(\left|\v{t}'\right|<\infty.\)
 
@@
 
+TODO: additional criteria are required for finite intensity process.
+
     We use only the left-continuous filtration
-\(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the paths of the process,
-which we here write as \(\cc{F}(N(t)).\)
+\(\mathcal{F}(\{N(s)\}_{s < t})\) generated by the paths of the process
+to time \(t\), which we save ink by writing here as
+\(\mathcal{F}(\{N(s)\}_{s < t}) \equiv \cc{H}_t.\)
 
-We call it a regular point process if it has an associated
-\emph{intensity} process \(\lambda\), such that \(\lambda\) is
-predictable with respect to \(\cc{F}(N)\) generated by the process, i.e.
-\cite{daley_introduction_2003}
+We call it a regular point process if we may characterise the times of
+the jumps of the process by an associated \emph{intensity} process
+\(\lambda\), such that \(\lambda\) is predictable with respect to
+\(\cc{H}_t\) i.e. \cite{daley_introduction_2003}
 
-\[\lambda(t|\cc{F}(N(t))):=\lim_{h\to 0} \frac{\bb E\left(N([t,t+h))\right)}{h}.\]
+\[\begin{aligned}\lambda(t|\cc{H}_t)&:=\lim_{h\to 0} \frac{1}{h}\left(1-\bb P\left[N(t+h)=N(t)|\cc{H}_t\right] \right)\\
+&=\lim_{h\to 0} \frac{1}{h}\left(\bb E \left[N([t,t+h))\right]\right).\end{aligned}\]
 
-Hereafter we define \(\lambda^*(t):=\lambda(t|\cc{F}(N(t)))\) for
-brevity.
+We will require that \(\bb E\lambda^*<\infty.\)
 
 Such regular point processes generalise the homogeneous Poisson rate in
-the sense that if we choose \(\lambda^*(t)\equiv \lambda_0,\) then it is
-the intensity of a homogeneous Poisson rate with fixed intensity
-\(\lambda_0\).
+the sense that if we choose constant
+\(\lambda(t|\cc{H}_t)\equiv \lambda_0,\) then it is the intensity of a
+homogeneous Poisson rate with fixed intensity \(\lambda_0\), such that a
+process \(N'\) with this intensity is Poisson,
+\(N'(t)-N'(s)\sim\oo{Pois}(\lambda_0(t-s))\) for \(t\geq s \geq 0\).
+More generally, we define \[
+\Lambda(s,t) = \int_s^t\lambda(u,\cc{H}_u) \dd u.
+\]
 
-    We will use the inter-arrival waiting times \(s_i=t_{i+1}-t_i\) to
-calculate the data likelihood, after Rubin \cite{rubin_regular_1972}.
+Then \[N(t)-N(s)\sim\oo{Pois}(\Lambda(s,t)),\] i.e. the pmf f
+\(k=N(t)-N(s)\) is given
 
-From the definition of the intensity we find \[\begin{aligned}
-\frac{\dd}{\dd t}\bb P(t_{i+1}=t|\mathcal{F}(N(t)))=
+\[\bb P(N(t)-N(s)=k)=\frac {\Lambda^k(s,t)e^{-\Lambda(s,t) }}{k!},\] and
+moreover, the increments of the process over disjoint intervals of the
+index space are independent.
+
+(TODO: not quite right - statement needs the ordering of the filtration
+to be true)
+
+    Considering each increment of the process , we have an expression for
+the density of a given arrival given the process history
+\[\begin{aligned}
+f(t, N(t_i)) &= \frac{\dd}{\dd t}\bb P(t_{i+1}<t) \\
+&= \frac{\dd}{\dd t} \bb P(N(t)=N(t_{i}))  \\
+&= \frac{\dd}{\dd t} \frac {\Lambda^0(t_i,t)e^{-\Lambda(t_i,t) }}{0!}\\
+&= \frac{\dd}{\dd t} e^{-\Lambda(t_i,t)} \\
+&= \left(\frac{\dd}{\dd t} -\Lambda(t_i,t)\right) e^{-\Lambda(t_i,t)} \\
+&= -\lambda(t, \cc H(t)) e^{-\Lambda(t_i,t)} 
 \end{aligned}\]
 
-TODO: Which bits does Z want proven? Rewrite to start from transition
-probability? Probably makes this shorter. Otherwise prove equivalence of
-intensity and transition prob DE?
+    then we may write the density of \(f(t_{i+1}),\)
+
+\[
+f(t_{i+1}, N(t_i)) = -\lambda(t_{i+1}, \cc H(t_{i+1})) e^{-\Lambda(t_i,t_{i+1})} 
+\] and the likelihood of the entire realisation as the product of all
+conditional densities
+
+\[
+f(t_1, t_2, \dots t_{N(T)}) = \prod_{i=0}^{N(T)} \left(-\lambda(t_{i+1}, \cc H(t_{i+1})) e^{-\Lambda(t_i,t_{i+1})} \right)
+\]
+
+TODO term for last interval \((t_{N(T)},T],\) handle \(i=0.\)
+
+    This is, as usual, more conveniently managed as a log likelihood
+
+\[\begin{aligned}
+\ell(\v t) &:= \log f(t_1, t_2, \dots t_{N(T)})\\
+& = \sum_{i=0}^{N(T)}\left( -\log \lambda(t_{i+1}, \cc H(t_{i+1})) -\Lambda(t_i,t_{i+1})\right)\\
+&=  \left(\sum_{i=1}^{N(T)} -\log \lambda(t_{i}, \cc H(t_{i}))\right) -\int_0^T \lambda(t, \cc H(t)
+\end{aligned}\]
+
+TODO missing minus sign for summation term.
 
@@
 \begin{aligned}
-\lambda(t|\mathcal{F}(N(t);\mu,\phi) &= \mu(t) + \phi * N\\
+\lambda(t|\cc H_t;\mu,\phi) &= \mu(t) + \phi * N\\
 &= \mu(t)  + \int_{-\infty}^{\infty}\phi(t-s)\dd N(s)\\
@@
 For brevity, we write
-\(\lambda(\mathcal{F}(N(t));\mu,\phi)=\lambda^*(t)\), but will suppress
-explicity dependence on history and parameters..
+\(\lambda(\mathcal{F}(N(t));\mu,\phi)=\lambda^*(t)\).
 
-    The log-likelihood function for a given realisation of the data is can
-be found from the usual likelihood for Poisson data of a given intensity
-\cite{ozaki_estimating_1979}
-
-\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda^*(t) \dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
-&=-\int_0^T\lambda^*(t)\dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
-&=-\int_0^T\lambda^*(t)\dd t + \sum_{t\in\v{t}'}\log \lambda^*(t)
+    Using that \(N(t)=N(t_i)\ \forall t\lt t_{i+1}\) we can now write the
+joint occurrence density as \cite{rubin_regular_1972}.
+\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda(t|\cc H_t) \dd t + \int_0^T\log \lambda(t|\cc H_t) \dd N_t \\ 
+    &=-\int_0^T\lambda(t|\cc H_t)\dd t + \int_0^T\log \lambda(t|\cc H_t) \dd N_t \\
+    &=-\int_0^T\lambda(t|\cc H_t)\dd t + \sum_{t\in\v{t}'}\log \lambda(t|\cc H_t) 
 \end{aligned}\]
 
-Using the method of maximum likelihood we estimate our parameter vector
-\(\theta:=(\mu, \phi)\) by choosing the estimates to maximise the
-log-likelihood,
-\[\hat{\theta}_\pi(N) = \oo{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
+    Substituting in the Hawkes intensity formula in particular gives us an
+expression for the joint likelihood.
+
+    Hereafter we define \(\lambda^*(t):=\lambda(t|\cc{H}_t)\) for brevity.
 
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_14_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_18_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_0.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_34_0.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_1.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_34_1.pdf}
     \end{center}
@@
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_30_2.pdf}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_34_2.pdf}
     \end{center}
@@
 
-\texttt{\color{outcolor}Out[{\color{outcolor}27}]:}
+\texttt{\color{outcolor}Out[{\color{outcolor}7}]:}
 
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_32_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_36_0.png}
     \end{center}
@@
 
-\texttt{\color{outcolor}Out[{\color{outcolor}28}]:}
+\texttt{\color{outcolor}Out[{\color{outcolor}8}]:}
 
     \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_34_0.png}
+    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{chapter_sparse_hawkes_files/chapter_sparse_hawkes_38_0.png}
     \end{center}
@@
 
+    This is suggestive of interpreting \(\lambda\) as a hazard rate, as
+indeed it is. Considering the density function \(f\) and cumulative
+density \(F\) of the time of the next event, TODO show this? start from
+this?
+
+\[\begin{aligned}
+\lambda(t)&=\frac{f(t|\cc{H}_t)}{1-F(t|\cc{H}_t)}\\
+&=\frac{f(t|\cc{H}_t)}{1-F(t|\cc{H}_t)}\\
+\end{aligned}\]
+
+    Rearranging the intensity, we note that \[\begin{aligned}
+\frac{\dd}{\dd t} \bb P\left[N(t)=N(s)|\cc{H}_t\right] &= \lambda(t|\cc{H}_t) \\
+&= 
+\end{aligned}\] for \(t\geq s \geq 0\)
+
+so that
+
+\[\begin{aligned}
+\bb P\left[N(t)=N(s)|\cc{H}_s\right]\propto\exp\left[-\int_{s}^{t}\lambda^*(u) \dd u\right]
+.\end{aligned}\]
+
     Given two ML fitted models, \((M_1,\hat{\theta}^M_1)\) and
@@
 
+    The log-likelihood function for a given realisation of the data is can
+be found from the usual likelihood for Poisson data of a given intensity
+\cite{ozaki_estimating_1979}
+
+\[\begin{aligned}L(N;\mu,\phi) &=-\int_0^T\lambda^*(t) \dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
+&=-\int_0^T\lambda^*(t)\dd t + \int_0^T\log \lambda^*(t) \dd N_t \\
+&=-\int_0^T\lambda^*(t)\dd t + \sum_{t\in\v{t}'}\log \lambda^*(t)
+\end{aligned}\]
+
+Using the method of maximum likelihood we estimate our parameter vector
+\(\theta:=(\mu, \phi)\) by choosing the estimates to maximise the
+log-likelihood,
+\[\hat{\theta}_\pi(N) = \oo{argmax}_{\mu,\phi} L_\pi(N;\theta).\]
+
+    \[\lambda(t|\cc{F}(N(t))):=\lim_{h\to 0} \frac{\bb E\left(N([t,t+h))\right)}{h}.\]
+
     % Add a bibliography block to the postdoc

### 2017-06-16 dodo.py
grad show pony

@@ -39,6 +39,7 @@ def _task_html(pth):
             # 'ln -f docs/refs.bib docs/'.format(pth=pth),
             'jupyter nbconvert --to html '
             # '--template=docs/{pth}_html.tpl '
+            '--TemplateExporter.exclude_output_prompt=True '
             '--FilesWriter.build_directory=docs/ '
             'docs/{pth}.ipynb'.format(pth=pth),
         ],
@@ -70,6 +71,7 @@ def _task_latex(pth):
             'ln -f docs/refs.bib _paper_output'.format(pth=pth),
             'jupyter nbconvert --to latex --template=docs/{pth}_print.tplx '
             '--FilesWriter.build_directory=_paper_output/ '
+            '--TemplateExporter.exclude_output_prompt=True '
             'docs/{pth}.ipynb'.format(pth=pth),
         ],
         clean=[
